### 2020-12-04 r_notebooks/exploring_oral_contraceptives.Rmd

---
title: "Oral contraceptives"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup}
## Packages
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)

## Options
knitr::opts_chunk$set(cache = TRUE)
options(mc.cores = parallel::detectCores() - 1)
rstan_options(auto_write = TRUE)
options(future.rng.onMisue = "ignore")

## Seed
set.seed(32)

## Functions
source("../r_scripts/functions.r")

## Drugs of interest
combined_hormonal <- 
  c("0703010E0", "0703010F0", "0703010G0", "0703010H0", "0703010M0",
    "0703010S0", "0703010R0","0703010A0", "0703010P0")
progestogen_only <- 
  c("0703021Q0", "0703022P0", "0703021C0", "0703021L0", "0703023L0",
    "0703022M0", "0703021N0", "0703022N0", "0703021P0")

oral_contraceptives <- c(combined_hormonal, progestogen_only)

## Read from database
# con <- DBI::dbConnect(RSQLite::SQLite(), "../OneDrive/1_data/gpextract/db-gpextract.sqlite")
# db.gpextract <- tbl(con, "gpextract")
# 
# oc <- db.gpextract %>%
#   filter(bnfchem %in% oral_contraceptives) %>%
#   collect() %>%
#   mutate(datename = dmy(paste0("28-", datename)))  # Add day component to date and convert to date-time object

## Write to local file
# oc %>%
#   write_rds("../data/oral_contraceptives.rds")

## Read from local file
gp <- read_rds("../data/gp_linked_na_omit.rds")
oc <- read_rds("../data/oral_contraceptives.rds")

## Sort into monthly count data
# oc_monthly_count <- oc %>% 
#   group_by(datename) %>% 
#   summarise(n = sum(Quantity)) %>% 
#   mutate(datename = yearmonth(datename),
#          n = case_when(str_detect(datename, pattern = "-01-") ~ n/31,
#                        str_detect(datename, pattern = "-02-") ~ n/28.25,
#                        str_detect(datename, pattern = "-03-") ~ n/31,
#                        str_detect(datename, pattern = "-04-") ~ n/30,
#                        str_detect(datename, pattern = "-05-") ~ n/31,
#                        str_detect(datename, pattern = "-06-") ~ n/30,
#                        str_detect(datename, pattern = "-07-") ~ n/31,
#                        str_detect(datename, pattern = "-08-") ~ n/31,
#                        str_detect(datename, pattern = "-09-") ~ n/30,
#                        str_detect(datename, pattern = "-10-") ~ n/31,
#                        str_detect(datename, pattern = "-11-") ~ n/30,
#                        str_detect(datename, pattern = "-12-") ~ n/31
#                              )) %>% 
#   ungroup() %>% 
#   as_tsibble(index = datename)
# 
# oc_monthly_quantity_locality <- 
#   right_join(oc, select(gp, c(PracticeID, npat, GPCluster)), 
#              by = c("PracticeID" = "PracticeID")) %>%
#   countRegional(., Locality, datename, Quantity)
# 
# oc_monthly_quantity_cluster <- 
#   right_join(oc, select(gp, c(PracticeID, npat, GPCluster)), 
#              by = c("PracticeID" = "PracticeID")) %>%
#   countRegional(., GPCluster, datename, Quantity) %>% 
#   mutate(n = 1000*n/npat) # quantity per 1000 pts
# 
# oc_monthly_quantity_gp <- 
#   countRegional(oc, PracticeID, datename, Quantity) %>% 
#   right_join(gp, by = c("regional_unit" = "PracticeID")) %>% # want to have GPs which have recorded data for the duration, not just part of it as that would add noise
#   mutate(n = 1000*n/npat) # quantity per 1000 pts
# 
# ## Sort into monthly count data pre-lockdown (2020-03-23)
# oc_monthly_count_pre_cv19 <- 
#   oc_monthly_count %>% 
#   filter(datename < dmy("01-03-2020"))
# 
# oc_monthly_quantity_locality_pre_cv19 <- 
#   oc_monthly_quantity_locality %>% 
#     filter(datename < dmy("01-03-2020"))
# 
# oc_monthly_quantity_cluster_pre_cv19 <- 
#   oc_monthly_quantity_cluster %>% 
#     filter(datename < dmy("01-03-2020"))
# 
# oc_monthly_quantity_gp_pre_cv19 <- 
#   oc_monthly_quantity_gp %>% 
#     filter(datename < dmy("01-03-2020"))

###
## Sort into monthly count data
oc_monthly_count <- oc %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(datename = yearmonth(datename),
         n = case_when(str_detect(datename, pattern = "Jan") ~ n/31,
                       str_detect(datename, pattern = "Feb") ~ n/28.25,
                       str_detect(datename, pattern = "Mar") ~ n/31,
                       str_detect(datename, pattern = "Apr") ~ n/30,
                       str_detect(datename, pattern = "May") ~ n/31,
                       str_detect(datename, pattern = "Jun") ~ n/30,
                       str_detect(datename, pattern = "Jul") ~ n/31,
                       str_detect(datename, pattern = "Aug") ~ n/31,
                       str_detect(datename, pattern = "Sep") ~ n/30,
                       str_detect(datename, pattern = "Oct") ~ n/31,
                       str_detect(datename, pattern = "Nov") ~ n/30,
                       str_detect(datename, pattern = "Dec") ~ n/31
                             )) %>% 
  ungroup() %>% 
  as_tsibble(index = datename)

oc_monthly_quantity_locality <- 
  right_join(oc, select(gp, c(PracticeID, npat, GPCluster)), 
             by = c("PracticeID" = "PracticeID")) %>%
  countRegional(., Locality, datename, Quantity)

oc_monthly_quantity_cluster <- 
  right_join(oc, select(gp, c(PracticeID, GPCluster)), 
             by = c("PracticeID" = "PracticeID")) %>%
  countRegional(., GPCluster, datename, Quantity) %>% 
  left_join(gp %>% 
  select(c(PracticeID, npat, GPCluster)) %>% 
  group_by(GPCluster) %>% 
  summarise(npat = sum(npat)), by = c("regional_unit" = "GPCluster")) %>% 
  mutate(n = 1000*n/npat) # quantity per 1000 pts

oc_monthly_quantity_gp <- 
  oc %>% 
  countRegional(., PracticeID, datename, Quantity) %>% 
  left_join(select(gp, c(PracticeID, npat)), 
            by = c("regional_unit" = "PracticeID")) %>% 
  mutate(n = 1000*n/npat)

## Sort into monthly count data pre-lockdown (2020-03-23)
oc_monthly_count_pre_cv19 <- 
  oc_monthly_count %>% 
  filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_locality_pre_cv19 <- 
  oc_monthly_quantity_locality %>% 
    filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_cluster_pre_cv19 <- 
  oc_monthly_quantity_cluster %>% 
    filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_gp_pre_cv19 <- 
  oc_monthly_quantity_gp %>% 
    filter(datename < dmy("01-03-2020"))

```

## Overview

### 2016-2020 plot of all oc prescriptions

```{r oc-monthly-ts-plot, fig.cap="Total quantity of combined hormonal and progestogen only oral contraceptives prescribed per month."}
oc %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  tsibble() %>% 
  autoplot(n) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

* General negative trend
* Seasonality unclear
  * Holiday effect? Drop in prescribing in Dec 2016-2019 followed by peak in Jan for 2017-2020.

### Seasonality

```{r oc-monthly-seasonality-plot}
df <- oc_monthly_count_pre_cv19 %>% 
  mutate(month = as.factor(month(datename, label = TRUE, abbr = TRUE)),
         year = as.factor(year(datename)))

df2 <- df %>% 
  as_tibble() %>% 
  filter(year(datename) != 2020) %>% 
  group_by(month) %>% 
  summarise(n = mean(n))

df %>% 
  ggplot(aes(x = month, y = n, colour = year, group = year)) +
  geom_line() +
  geom_line(data = df2, aes(x = month, y = n, group = "Avg"), 
            colour = "red", lty = 2) +
  labs(caption = "Red dashed line is mean for 2016-2019")

```
* 2016 is very different. Is that grounds to remove it?
* Dec/Jan holiday effect
  * How, if at all, does the population of Wales change with students?
* Easter effect? Mar/Apr has a pattern with 2017 as exception (does the opposite)
  * What would be expected? If people are travelling for Easter, then they may get their prescriptions before and after, but not in the weeks around?
  * Range (2000-2020): March 23rd (2008) to April 21st (2019)
  * Dates:
    * 2016-03-27
    * 2017-04-16
    * 2018-04-01
    * 2019-04-21

    * 2020-04-12

```{r}
oc_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename)) %>% 
  gg_subseries(y = n, period = "year")
```

*

### By `Quantity`
* Jan high and Dec low

### By `Items`
* Feb is lower, but has fewer days in it
* Dec is lower, holiday effect?
* Jan is higher, people who haven't had prescriptions due to holidays/have been entered late onto the system due to holidays

### Lag plots

```{r}
oc_monthly_count_pre_cv19 %>%
  mutate(datename = yearmonth(datename)) %>% 
  gg_lag(y = n, geom = "point", lags = 1:12)
```

```{r}
oc_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename)) %>% 
  gg_lag(y = n, geom = "point", lags = 13:24)
```
* Here the colours indicate the month of the variable on the y-axis. The relationship is strongest at lag 12 and 24 suggesting seasonal data

### Autocorrelation

```{r}
tseries::adf.test(oc_monthly_count$n)
urca::ur.df(oc_monthly_count$n)
```

```{r}
oc_monthly_count_pre_cv19 %>%
  mutate(datename = yearmonth(datename)) %>% 
  ACF(n, lag_max = 60) %>% autoplot()
```

* Autocorrelation greatest at 2 and 12

```{r}
oc_monthly_count_pre_cv19$n %>%
  pacf(lag.max = 60)
```

## Time series components

```{r}
dcmp <- 
  oc_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename)) %>% 
  model(STL(n))

oc_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename)) %>% 
  autoplot(n, color = "grey") + 
  autolayer(components(dcmp), trend, color = "red")

components(dcmp) %>% 
  autoplot()
```

```{r}
oc_monthly_count %>% 
  mutate(datename = yearmonth(datename)) %>% 
  autoplot(n, color = "grey") + 
  autolayer(components(dcmp), season_adjust, colour = "blue") +
  labs(title = "Seasonally adjusted")
```

## Moving average

```{r}
df <- oc_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename),
         `5-MA` = slider::slide_dbl(n, mean, .before = 2, .after = 2, 
                                    .complete = TRUE),
         `12-MA` = slider::slide_dbl(n, mean, .before = 5, .after = 6, 
                                    .complete = TRUE),
         `2x12-MA` = slider::slide_dbl(`12-MA`, mean, .before = 2, .after = 2, 
                                    .complete = TRUE),
         )
df %>% 
  autoplot(n) + 
  autolayer(df, `5-MA`, colour = "red") +
  autolayer(df, `2x12-MA`, colour = "blue") +
  labs(caption = "Daily prescriptions (black) along with the 5-MA estimate of the trend-cycle (red) and 2x12-MA (blue).")
```

## Predictions by `Locality`

```{r}
regions <- unique(as.character(oc_monthly_quantity_locality$regional_unit))
holidays <- 
  bind_rows(tibble(holiday = "christmas",
                   ds = as.Date(c("2016-12-25", "2017-12-25", "2018-12-25",
                                  "2019-12-25", "2020-12-25", "2021-12-25")),
                   lower_window = -1, 
                   upper_window = 1),

            tibble(holiday = "easter",
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -1, 
                   upper_window = 1))

fit <- oc_monthly_quantity_locality_pre_cv19 %>% 
  # filter(year(datename) > 2016) %>% 
  mutate(datename = yearmonth(datename)) %>% 
  model(#trend_model1 = TSLM(n ~ trend()),#time series linear model
        trend_model2 = TSLM(n ~ trend() + season()),
        #ets1 = ETS(n ~ trend()),
        ets2 = ETS(n ~ trend() + season()),
        arima = ARIMA(n),# stepwise = FALSE, approximation = FALSE),
        #comb1 = combination_model(TSLM(n ~ trend()), ETS(n ~ trend())),
        comb2 = combination_model(TSLM(n ~ trend() + season()), 
                                  ETS(n ~ trend() + season())),
        # stl_dcmp1 = decomposition_model(STL(n ~ trend(),
        #                                    iterations = 1000, robust = FALSE),
        #                                NAIVE(season_adjust)),
        stl_dcmp2 = decomposition_model(STL(n ~ trend() + season(),
                                            iterations = 1000, robust = FALSE),
        SNAIVE(season_adjust)),
        s_naive = SNAIVE(n)
  )

fcsts <-  fit %>% 
  forecast(h = "6 months", bootstrap = FALSE) # can use `times = x` to control the number of samples

# single thread
regional_plots <- 
  map(.x = regions, 
      .f = ~regionalPlot(.x, oc_monthly_quantity_locality))

regional_plots

left_join(accuracy(fit),
          fit %>% 
            select(-starts_with(c("stl_dcmp", "comb"))) %>% 
            report(), 
by = c("regional_unit", ".model"))

fit %>% 
  accuracy() %>% 
  group_by(.model) %>% 
  summarise(across(is.numeric, .fns = ~mean(., na.rm = TRUE)))

```

* https://fabletools.tidyverts.org/reference/combination_model.html

### Residual diagnostics

```{r}
augment(fit) %>% 
  features(.resid, box_pierce, lag = 24, dof = 0) %>% 
  mutate(sig = if_else(bp_pvalue <.05, TRUE, FALSE))
```

## Predictions by `GPCluster`

```{r}
# system.time(cluster_fit <- regionalJointFits(oc_monthly_quantity_cluster_pre_cv19))

```

### Residual diagnostics

```{r}
augment(fit) %>% 
  features(.resid, box_pierce, lag = 24, dof = 0) %>% 
  mutate(sig = if_else(bp_pvalue <.05, TRUE, FALSE))
```

## Predictions by `PracticeID`

```{r}
regions <- unique(as.character(gp$PracticeID))
holidays <- 
  bind_rows(tibble(holiday = "christmas",
                   ds = as.Date(c("2016-12-25", "2017-12-25", "2018-12-25",
                                  "2019-12-25", "2020-12-25", "2021-12-25")),
                   lower_window = -1, 
                   upper_window = 1),

            tibble(holiday = "easter",
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -1, 
                   upper_window = 1))

fit <- oc_monthly_quantity_gp_pre_cv19 %>% 
    filter(regional_unit %in% gp$PracticeID) %>% 
  mutate(datename = yearmonth(datename)) %>% 
  model(#trend_model1 = TSLM(n ~ trend()),#time series linear model
        trend_model2 = TSLM(n ~ trend() + season()),
        ets2 = ETS(n ~ trend() + season()),
        arima = ARIMA(n),# stepwise = FALSE, approximation = FALSE),

        comb2 = combination_model(TSLM(n ~ trend() + season()), 
                                  ETS(n ~ trend() + season())),
        stl_dcmp2 = decomposition_model(STL(n ~ trend() + season(),
                                            iterations = 1000, robust = FALSE),
        SNAIVE(season_adjust)),
        # s_naive = SNAIVE(n)
  )

fcsts <-  fit %>% 
  forecast(h = "5 months", bootstrap = FALSE) # can use `times = x` to control the number of samples

# single thread
regional_plots <- 
  map(.x = regions, 
      .f = ~regionalPlot(.x, filter(oc_monthly_quantity_gp, 
                                    regional_unit %in% gp$PracticeID)))

regional_plots

left_join(accuracy(fit),
          fit %>% 
            select(-starts_with(c("stl_dcmp", "comb"))) %>% 
            report(), 
by = c("regional_unit", ".model"))

fit %>% 
  accuracy() %>% 
  group_by(.model) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

```

### Residual diagnostics

```{r}
augment(fit) %>% 
  features(.resid, box_pierce, lag = 24, dof = 0) %>% 
  mutate(sig = if_else(bp_pvalue <.05, TRUE, FALSE))
```

## Overview of each drug

```{r}
oc %>% 
  group_by(datename, chemname) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(n = case_when(str_detect(datename, pattern = "-01-") ~ n/31,
                       str_detect(datename, pattern = "-02-") ~ n/28.25,
                       str_detect(datename, pattern = "-03-") ~ n/31,
                       str_detect(datename, pattern = "-04-") ~ n/30,
                       str_detect(datename, pattern = "-05-") ~ n/31,
                       str_detect(datename, pattern = "-06-") ~ n/30,
                       str_detect(datename, pattern = "-07-") ~ n/31,
                       str_detect(datename, pattern = "-08-") ~ n/31,
                       str_detect(datename, pattern = "-09-") ~ n/30,
                       str_detect(datename, pattern = "-10-") ~ n/31,
                       str_detect(datename, pattern = "-11-") ~ n/30,
                       str_detect(datename, pattern = "-12-") ~ n/31
                             )) %>% 
  as_tsibble(index = datename, key = chemname) %>% 
  autoplot(n) +
  theme(legend.position = "none") +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

## Modelling

### By GP Cluster

```{r}
# system.time(oc_gp_cluster_cv_models <-
#   regionalCV(data = oc_monthly_quantity_cluster_pre_cv19))
# 
# oc_gp_cluster_cv_models %>%
#   write_rds("../data/oc_gp_cluster_cv_models.rds")

oc_gp_cluster_cv_models <- 
  read_rds("../data/oc_gp_cluster_cv_models.rds")

oc_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = MAPE)) +
  geom_histogram()

oc_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))

oc_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  ggplot(aes(x = reorder(.model, MAPE, mean), y = MAPE)) + # reorder by mean of y
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
system.time(oc_gp_cluster_fcts <- regionalJointFcsts(oc_monthly_quantity_cluster_pre_cv19))

oc_gp_cluster_fcts
```

## Discrepancies

```{r}
best_model <- 
  oc_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_cluster_fcts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd) %>% 
  ggplot(aes(y = discrepancy)) +
  # geom_boxplot(aes(x = as.factor(datename))) +
  geom_line(aes(x = datename, y = discrepancy, colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none") #+
#  coord_cartesian(ylim = c(-1,1)) +
  facet_wrap(~.model)
```

```{r}
library(tidyLPA)
best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_cluster_fcts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd)  %>%
    select(discrepancy, regional_unit, datename) %>%
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  ungroup() %>% 
  select(-regional_unit) %>% 
  estimate_profiles(5) %>% plot_profiles()
    estimate_profiles(1:5,
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")
                      ) %>% 
    compare_solutions(statistics = c("AIC", "BIC"))
```

### Plotting

```{r}
best_model <- 
  oc_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oc_gp_cluster_fcts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oc_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = .mean, colour = as.factor(regional_unit))) +
  geom_ribbon(aes(ymin = lb_80, ymax = ub_80, fill = as.factor(regional_unit),
              alpha = .05), outline.type = "both") +
  geom_line(aes(y = n, group = regional_unit, colour = as.factor(regional_unit)), lty = 3) +
  coord_cartesian(ylim = c(0,50))+
  theme(legend.position = "none") +
  facet_wrap(~.model) + 
  labs(caption = "Dotted line = observed data, solid line and ribbon is predicted data with 80% CIs")

```

```{r}
bestModelPlot <- function(models, forecasts, data, region){
  best_model <- 
    models %>% 
    group_by(regional_unit) %>% 
    arrange(RMSE, MAE) %>% 
    slice(1, .preserve = TRUE)

  full_data <- filter(data, 
                      datename > ymd("2020-02-28") &
                        regional_unit == region) %>% 
    mutate(datename = yearmonth(datename))

best_model %>% 
    select(.model, regional_unit) %>% 
    filter(regional_unit == region) %>% 
    inner_join(forecasts, by = c(".model", "regional_unit")) %>%
    left_join(full_data, by = c("datename", "regional_unit")) %>% 
    ggplot(aes(x = datename, colour = as.factor(regional_unit))) +
    geom_line(aes(y = .mean)) +
    geom_ribbon(aes(ymin = lb_80, ymax = ub_80, fill = as.factor(regional_unit),
                   alpha = .1)) +
    geom_line(aes(y = n),
              colour = "black") +
    theme(legend.position = "none") +
    facet_wrap(~.model)
}

bestModelPlot(models = oc_gp_cluster_cv_models, 
              forecasts = oc_gp_cluster_fcts$forecasts,
              data = oc_monthly_quantity_cluster,
              region = "CityHealth")

```

## Other
### Prophet

```{r}
# Prophet model
oc_prophet_model <- 
  fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                           season("year", type = "additive") + holiday(holidays))

fit <- oc_monthly_quantity_locality_pre_cv19 %>% 
  as_tsibble(index = datename) %>% 
  model(oc_prophet_model)

fc <- 
  fit %>% 
  forecast(h = "5 months")

accuracy(fit)

df <- oc_monthly_count_pre_cv19 %>% 
  transmute(ds = datename, y = n)
m <- prophet::prophet(df, seasonality.mode = "additive")
future <- make_future_dataframe(m, periods = 6, freq = 'month')
fcst <- predict(m, future)
prophet_plot_components(m, fcst)

oc_monthly_count %>% 
  ggplot(aes(x = datename, y = n)) +
  geom_line() +
  geom_ribbon(aes(ymin = c(rep(NA,50), fcst[51:55,]$yhat_lower),
                  ymax = c(rep(NA,50), fcst[51:55,]$yhat_upper)),
              alpha = .1, fill = "blue") +
  geom_line(aes(y = c(fcst[1:50,]$yhat, rep(NA, 5))),
            col = "blue", lty = 2) +
  geom_line(aes(y = c(rep(NA,50), fcst[51:55,]$yhat)),
            col = "blue", lty = 3) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown

perf <- cross_validation(model = m, initial = 3*52, horizon = 52, units = "weeks")
performance_metrics(perf)
plot_cross_validation_metric(perf, metric = "rmse")
```

## Split by practice

```{r}
oc %>% 
  tibble %>% 
  group_by(datename, PracticeID) %>% 
  count %>% 
  group_by(PracticeID) %>% 
  summarise(n_min = min(n), 
            n_mean = mean(n)) %>% arrange((n_min)) %>% 
  filter(n_min >1)

oc %>% filter(PracticeID == "W91626") %>% 
  group_by(datename) %>% 
  count() %>% 
  as_tsibble(index = datename) %>% 
  autoplot()

```