### 2020-12-11 r_notebooks/exploring_oral_anticoagulants.Rmd
Updataed gp data analyses

---
title: "Oral anticoagulants"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup}
## Packages
# devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(serCymruTools)

## Options
knitr::opts_chunk$set(cache = TRUE)
options(mc.cores = parallel::detectCores() - 1)
rstan_options(auto_write = TRUE)

## Seed
set.seed(32)

```

## Preparing data

```{r prep-data}
## Drugs of interest
oral_anticoagulants <- 
  c("0208020H0", "0208020Z0", "0208020X0", "0208020AA", "0208020W0", 
    "020802000", "0208020I0", "0208020N0", "0208020S0", "0208020Y0", 
    "0208020V0")

# ## Read from database
# con <- DBI::dbConnect(RSQLite::SQLite(), 
#                       "../OneDrive/1_data/gpextract/db-gpextract.sqlite")
# db.gpextract <- tbl(con, "gpextract")
# oac <- db.gpextract %>%
#   filter(bnfchem %in% oral_anticoagulants) %>%
#   collect() %>%
#   mutate(datename = dmy(paste0("28-", datename)))  # Add day component to date and convert to date-time object
# DBI::dbDisconnect(con)
# 
# ## Write to local file
# oac %>%
#   write_rds("../data/oral_anticoagulants.rds")

## Read from local file
gp <- read_rds("../OneDrive/2_interimdata/8_linkeddata/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL) %>% 
  filter(year == 2020) # do something better
oac <- read_rds("../data/oral_anticoagulants.rds")

## Sort into monthly count data
oac_monthly_count <- oac %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(datename = yearmonth(datename),
         n = case_when(str_detect(datename, pattern = "Jan") ~ n/31,
                       str_detect(datename, pattern = "Feb") ~ n/28.25,
                       str_detect(datename, pattern = "Mar") ~ n/31,
                       str_detect(datename, pattern = "Apr") ~ n/30,
                       str_detect(datename, pattern = "May") ~ n/31,
                       str_detect(datename, pattern = "Jun") ~ n/30,
                       str_detect(datename, pattern = "Jul") ~ n/31,
                       str_detect(datename, pattern = "Aug") ~ n/31,
                       str_detect(datename, pattern = "Sep") ~ n/30,
                       str_detect(datename, pattern = "Oct") ~ n/31,
                       str_detect(datename, pattern = "Nov") ~ n/30,
                       str_detect(datename, pattern = "Dec") ~ n/31)) %>% 
  ungroup() %>% 
  as_tsibble(index = datename)

oac_monthly_quantity_hb <- 
  right_join(oac, select(gp, c(practiceid, hboard)), 
             by = c("PracticeID" = "practiceid")) %>%
  countRegional(hboard, datename, Quantity) %>% 
  left_join(gp %>% 
  select(c(practiceid, npat, hboard)) %>% 
  group_by(hboard) %>% 
  summarise(npat = sum(npat)), by = c("regional_unit" = "hboard")) %>% 
  mutate(n = 1000*n/npat) # quantity per 1000 pts

oac_monthly_quantity_cluster <- 
  right_join(oac, select(gp, c(practiceid, cluster)), 
             by = c("PracticeID" = "practiceid")) %>%
  countRegional(cluster, datename, Quantity) %>% 
  left_join(gp %>% 
  select(c(practiceid, npat, cluster)) %>% 
  group_by(cluster) %>% 
  summarise(npat = sum(npat)), by = c("regional_unit" = "cluster")) %>% 
  mutate(n = 1000*n/npat) # quantity per 1000 pts

oac_monthly_quantity_gp <- 
  countRegional(oac, PracticeID, datename, Quantity) %>% 
  right_join(select(gp, c(practiceid, npat)), 
            by = c("regional_unit" = "practiceid")) %>% 
  mutate(n = 1000*n/npat)

## Sort into monthly count data pre-lockdown (2020-03-23)
oac_monthly_count_pre_cv19 <- 
  oac_monthly_count %>% 
  filter(datename < dmy("01-03-2020"))

oac_monthly_quantity_hb_pre_cv19 <- 
  oac_monthly_quantity_hb %>% 
    filter(datename < dmy("01-03-2020"))

oac_monthly_quantity_cluster_pre_cv19 <- 
  oac_monthly_quantity_cluster %>% 
    filter(datename < dmy("01-03-2020"))

oac_monthly_quantity_gp_pre_cv19 <- 
  oac_monthly_quantity_gp %>% 
    filter(datename < dmy("01-03-2020"))

```

## Overview

### 2016-2020 plot of all oac prescriptions

```{r oac-monthly-ts-plot, fig.cap="Total quantity of combined oral anticoagulants prescribed per month."}
oac %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  tsibble() %>% 
  autoplot(n) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

### Seasonality

```{r oac-monthly-seasonality-plot}
df <- oac_monthly_count_pre_cv19 %>% 
  mutate(month = as.factor(month(datename, label = TRUE, abbr = TRUE)),
         year = as.factor(year(datename)))

df2 <- df %>% 
  as_tibble() %>% 
  filter(year(datename) != 2020) %>% 
  group_by(month) %>% 
  summarise(n = mean(n))

df %>% 
  ggplot(aes(x = month, y = n, colour = year, group = year)) +
  geom_line() +
  geom_line(data = df2, aes(x = month, y = n, group = "Avg"), 
            colour = "red", lty = 2) +
  labs(caption = "Red dashed line is mean for 2016-2019")

```
*

```{r}
oac_monthly_count_pre_cv19 %>% 
  mutate(datename = yearmonth(datename)) %>% 
  gg_subseries(y = n, period = "year")
```

### Autocorrelation

```{r}
tseries::adf.test(oac_monthly_count$n)
urca::ur.df(oac_monthly_count$n)
```

```{r}
oac_monthly_count_pre_cv19 %>%
  mutate(datename = yearmonth(datename)) %>% 
  ACF(n, lag_max = 24) %>% autoplot()
```

* Autocorrelation greatest at 2 and 12

```{r}
oac_monthly_count_pre_cv19$n %>%
  pacf(lag.max = 24)
```

## Modelling

### By Health Board

```{r, health-board-modelling}
# system.time(oac_hb_cv_models <-
#   regionalCV(data = oac_monthly_quantity_hb_pre_cv19))
# 
# oac_hb_cv_models %>%
#   write_rds("../data/oac_hb_cv_models.rds")

oac_hb_cv_models <- 
  read_rds("../data/oac_hb_cv_models.rds")

oac_hb_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = MAPE)) +
  geom_histogram()

oac_hb_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))

oac_hb_cv_models %>% 
  group_by(regional_unit) %>% 
  ggplot(aes(x = reorder(.model, MAPE, mean), y = MAPE)) + # reorder by mean of y
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))
```

* 724.044 seconds with multisession prophetModels()

```{r}
# system.time(oac_hb_fcsts <- regionalJointFcsts(oac_monthly_quantity_hb_pre_cv19))
# 
# oac_hb_fcsts %>%
#   write_rds("../data/oac_hb_fcsts.rds")

oac_hb_fcsts <- 
  read_rds("../data/oac_hb_fcsts.rds")
```

* 194.071 seconds 2020-12-10

### By GP Cluster

```{r, gp-cluster-modelling}
# system.time(oac_gp_cluster_cv_models <-
#   regionalCV(data = oac_monthly_quantity_cluster_pre_cv19))
# 
# oac_gp_cluster_cv_models %>%
#   write_rds("../data/oac_gp_cluster_cv_models.rds")

oac_gp_cluster_cv_models <- 
  read_rds("../data/oac_gp_cluster_cv_models.rds")

oac_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = MAPE)) +
  geom_histogram()

oac_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))

oac_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  ggplot(aes(x = reorder(.model, MAPE, mean), y = MAPE)) + # reorder by mean of y
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))

```

* 5514.279 seconds

```{r}
# system.time(oac_gp_cluster_fcsts <- regionalJointFcsts(oac_monthly_quantity_cluster_pre_cv19))
# 
# oac_gp_cluster_fcsts %>%
#   write_rds("../data/oac_gp_cluster_fcsts.rds")

oac_gp_cluster_fcsts <- 
  read_rds("../data/oac_gp_cluster_fcsts.rds")
```

* 1266.126 seconds 2020-12-10

### By Practice

```{r, gp-practice-modelling}
incomplete_gp <- oac_monthly_quantity_gp_pre_cv19 %>% 
    as_tibble() %>% 
    group_by(regional_unit) %>% 
    count() %>% 
    filter(n < 50) %>% 
    .$regional_unit %>% 
    unique()

# system.time(oac_gp_cv_models <-
#   oac_monthly_quantity_gp_pre_cv19 %>%
#     filter(!regional_unit %in% incomplete_gp) %>%
#               regionalCV())
# 
# oac_gp_cv_models %>%
#   write_rds("../data/oac_gp_cv_models.rds")

oac_gp_cv_models <- 
  read_rds("../data/oac_gp_cv_models.rds")

oac_gp_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = MAPE)) +
  geom_histogram()

oac_gp_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))

oac_gp_cv_models %>% 
  group_by(regional_unit) %>% 
  ggplot(aes(x = reorder(.model, MAPE, mean), y = MAPE)) + # reorder by mean of y
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90))

```

* 2.195e+04 seconds 2020-12-07
* 25547.266 seconds 2020-12-09

```{r}
system.time(oac_gp_fcsts <-
              oac_monthly_quantity_gp_pre_cv19 %>%
              filter(!regional_unit %in% incomplete_gp) %>%
              regionalJointFcsts())
oac_gp_fcsts %>%
  write_rds("../data/oac_gp_fcsts.rds")

oac_gp_cluster_fcsts <- 
  read_rds("../data/oac_gp_fcsts.rds")
```

## Discrepancies

### By Health Board

```{r, health-board-discrepancies}
best_model_hb <- 
  oac_hb_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model_hb %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oac_hb_fcsts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oac_monthly_quantity_hb, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd) %>% 
  ggplot(aes(y = discrepancy)) +
  geom_line(aes(x = datename, y = discrepancy, colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none")
```

### By GP Cluster

```{r, gp-cluster-discrepancies}
best_model_gp_cluster <- 
  oac_gp_cluster_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model_gp_cluster %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oac_gp_cluster_fcsts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oac_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd) %>% 
  ggplot(aes(y = discrepancy)) +
  # geom_boxplot(aes(x = as.factor(datename))) +
  geom_line(aes(x = datename, y = discrepancy, colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  # scale_color_gradientn(colours = rainbow(100))+
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none")
```

```{r}
library(tidyLPA)
best_model_gp_cluster %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oac_gp_cluster_fcsts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oac_monthly_quantity_cluster, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd)  %>%
    select(discrepancy, regional_unit, datename) %>%
  pivot_wider(names_from = datename, values_from = discrepancy) %>% 
  ungroup() %>% 
  select(-regional_unit) %>% 
  # estimate_profiles(n_profiles = 4, models = 1) %>% plot_profiles()
    estimate_profiles(1:5,
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")
                      ) %>% 
    compare_solutions(statistics = c("AIC", "BIC"))
```

### By GP Practice

```{r, gp-practice-discrepancies}
best_model_gp <- 
  oac_gp_cv_models %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

best_model_gp %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(oac_gp_fcsts$forecasts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(oac_monthly_quantity_gp, 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - .mean,
         z_discrepancy = discrepancy/sd) %>% 
  ggplot(aes(y = discrepancy)) +
  # geom_boxplot(aes(x = as.factor(datename))) +
  geom_line(aes(x = datename, y = discrepancy, colour = as.factor(regional_unit)),
            alpha = .8, lty = 2) +
  geom_hline(aes(yintercept = 0), colour = "red") +
  theme(legend.position = "none")
```