### 2021-01-08 r_notebooks/parallel_forecasting.Rmd
Addibg mplus plotting functions from http://www.statmodel.com/mplus-R/

---
title: "Parallel forecasting"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, setup}
# Packages ####
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(knitr)
library(serCymruTools)

# Options ####
knitr::opts_chunk$set(cache = TRUE)
options(mc.cores = parallel::detectCores() - 1)
rstan_options(auto_write = TRUE)
options(future.rng.onMisue = "ignore")

# Seed ####
set.seed(32)

# Data ####

## Read from local file
#gp <- read_rds("../data/gp_linked_na_omit.rds")
gp <- read_rds("../OneDrive/2_interimdata/8_linkeddata/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL) %>% 
  filter(year == 2020) # do something better
oc <- read_rds("../data/oral_contraceptives.rds")

## Sort into monthly count data
oc_monthly_count <- oc %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(datename = yearmonth(datename),
         n = case_when(str_detect(datename, pattern = "Jan") ~ n/31,
                       str_detect(datename, pattern = "Feb") ~ n/28.25,
                       str_detect(datename, pattern = "Mar") ~ n/31,
                       str_detect(datename, pattern = "Apr") ~ n/30,
                       str_detect(datename, pattern = "May") ~ n/31,
                       str_detect(datename, pattern = "Jun") ~ n/30,
                       str_detect(datename, pattern = "Jul") ~ n/31,
                       str_detect(datename, pattern = "Aug") ~ n/31,
                       str_detect(datename, pattern = "Sep") ~ n/30,
                       str_detect(datename, pattern = "Oct") ~ n/31,
                       str_detect(datename, pattern = "Nov") ~ n/30,
                       str_detect(datename, pattern = "Dec") ~ n/31)) %>% 
  ungroup() %>% 
  as_tsibble(index = datename)
colnames(gp)

oc_monthly_quantity_hb <- 
  right_join(oc, select(gp, c(practiceid, hboard)), 
             by = c("PracticeID" = "practiceid")) %>%
  countRegional(hboard, datename, Quantity) %>% 
  left_join(gp %>% 
  select(c(practiceid, npat, hboard)) %>% 
  group_by(hboard) %>% 
  summarise(npat = sum(npat)), by = c("regional_unit" = "hboard")) %>% 
  mutate(n = 1000*n/npat) # quantity per 1000 pts

oc_monthly_quantity_cluster <- 
  right_join(oc, select(gp, c(practiceid, cluster)), 
             by = c("PracticeID" = "practiceid")) %>%
  countRegional(cluster, datename, Quantity) %>% 
  left_join(gp %>% 
  select(c(practiceid, npat, cluster)) %>% 
  group_by(cluster) %>% 
  summarise(npat = sum(npat)), by = c("regional_unit" = "cluster")) %>% 
  mutate(n = 1000*n/npat) # quantity per 1000 pts

oc_monthly_quantity_gp <- 
  countRegional(oc, PracticeID, datename, Quantity) %>% 
  right_join(select(gp, c(practiceid, npat)), 
            by = c("regional_unit" = "practiceid")) %>% 
  mutate(n = 1000*n/npat)

## Sort into monthly count data pre-lockdown (2020-03-23)
oc_monthly_count_pre_cv19 <- 
  oc_monthly_count %>% 
  filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_hb_pre_cv19 <- 
  oc_monthly_quantity_hb %>% 
    filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_cluster_pre_cv19 <- 
  oc_monthly_quantity_cluster %>% 
    filter(datename < dmy("01-03-2020"))

oc_monthly_quantity_gp_pre_cv19 <- 
  oc_monthly_quantity_gp %>% 
    filter(datename < dmy("01-03-2020"))

## Smaller GP data set to develop process
regions <- unique(as.character(gp$practiceid))
dev_data <- 
  oc_monthly_quantity_gp_pre_cv19 %>% 
  filter(regional_unit %in% sample(regions, 5))

rm(oc)
rm(gp)
```

```{r}
dev_data %>% 
  fableModels()
```

## Fable fits

* Fits can be estimated across several different fable models
* This can be done on the original data frame, or on each practice individually
* Need to ensure that models are the same using each method

### As single dataframe

```{r}
# As one df
system.time(fit <- dev_data %>% 
  filter(regional_unit %in% gp$PracticeID) %>% 
  mutate(datename = yearmonth(datename)) %>% 
  fableModels())

fits1 <- accuracy(fit) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

fits1 %>% 
  kable(digits = 2)
```

### map()

#### Single core

* `model()` takes advantage of `plan(multisession)` to automatically compute in parallel, but does not include the proper options (i.e., `furrr_options(seed = TRUE)`) and generates a warning. Does not appear to influence estimates.

```{r, eval=FALSE}
system.time(fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFableFits(dev_data, .x)))

fits2 <- accuracy(fit2) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

fits2 %>% 
  kable(digits = 2)
```

#### Multi core

```{r}
plan(multisession)

system.time(
fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFableFits(dev_data, .x), 
             .options = furrr_options(seed = TRUE)))

fits3 <- accuracy(fit3) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

fits3 %>% 
  kable(digits = 2)

plan(sequential)
```

### Are the results the same?

```{r}
fits1 == fits3

head(fits1); head(fits3)
```

* Yes

### Function to provide model

## Prophet fits

* Prophet is fastest as a single data frame, but are its predictions unique

### As single dataframe

```{r}
oc_prophet_model1 <- 
  fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                           season("year", type = "additive") + holiday(holidays))
oc_prophet_model2 <- 
  fable.prophet::prophet(n ~ growth("linear",) +
                           season("year", type = "additive") + holiday(holidays))
oc_prophet_model3 <- 
  fable.prophet::prophet(n ~ season("year", type = "additive") + holiday(holidays))
oc_prophet_model4 <- 
  fable.prophet::prophet(n ~ holiday(holidays))
oc_prophet_model5 <- 
  fable.prophet::prophet(n)

system.time(proph_fit <- dev_data %>% 
              mutate(datename = yearmonth(datename)) %>% 
              as_tsibble(index = datename) %>% 
              model(oc_prophet_model1, oc_prophet_model2, oc_prophet_model3, 
                    oc_prophet_model4, oc_prophet_model5))

proph_fits <- accuracy(proph_fit) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

proph_fits %>% 
  kable(digits = 2)

```

### map()

#### Single core
##### Sequential

```{r}
system.time(proph_fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalProphetFits(dev_data, .x)))

proph_fits2 <- accuracy(proph_fit2) %>% 
  group_by(.model) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

proph_fits2 %>% 
  kable(digits = 2)
```

### future_map()

##### multisession

```{r, eval=FALSE}
plan(multisession)

system.time(proph_fit2 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalProphetFits(dev_data, .x),
            .options = furrr_options(seed = TRUE)))

proph_fits2 <- accuracy(proph_fit2) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

proph_fits2 %>% 
  kable(digits = 2)
plan(sequential)

```

* 2020-12-07 Fails:

```
1 error encountered for oc_prophet_model1
[1] unable to load shared object '/Library/Frameworks/R.framework/Versions/4.0/Resources/library/prophet/libs/prophet.so':
  dlopen(/Library/Frameworks/R.framework/Versions/4.0/Resources/library/prophet/libs/prophet.so, 6): Library not loaded: @rpath/libtbb.dylib
  Referenced from: /Library/Frameworks/R.framework/Versions/4.0/Resources/library/prophet/libs/prophet.so
  Reason: image not found
```

* 2020-12-08 Fixed
  * Reinstalling using `install.packages("prophet", type="source")` has resolved problem.
  * Need to test in Windows

#### Multi core

```{r, eval=FALSE}
plan(multisession)
system.time(
  proph_fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
                               .f = ~regionalProphetFits(dev_data, .x), 
                               .options = furrr_options(seed = TRUE)))

proph_fits3 <- accuracy(proph_fit3) %>% 
  group_by(regional_unit) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE)))

proph_fits3 %>% 
  kable(digits = 2)

plan(sequential)
```

## Model comparions

```{r}
bind_rows(accuracy(fit3), accuracy(proph_fit), .id = "a") %>% 
  group_by(.model, a) %>% 
  summarise(across(where(is.numeric), .fns = ~mean(., na.rm = TRUE))) %>% 
  arrange(RMSE) %>% 
  kable(digits = 2)
```
### Plots

## Joint fable and prophet fits

```{r}

```

## LOO CV

### Example

From https://otexts.com/fpp3/tscv.html

```{r}
## example 
google_2015 <- 
  tsibbledata::gafa_stock %>% 
  filter(Symbol == "GOOG") %>% 
  fill_gaps()

# Time series cross-validation accuracy
google_2015_tr <- google_2015 %>%
  slice(1:(n()-1)) %>%
  stretch_tsibble(.init = 3, .step = 1)
fc <- google_2015_tr %>%
  model(RW(Close ~ drift())) %>%
  forecast(h = 1)

fc %>% accuracy(google_2015)

# Residual accuracy
google_2015 %>%
  model(RW(Close ~ drift())) %>%
  accuracy()
```

#### Plotting RMSE for different forecast windows

```{r}
google_2015_tr <- google_2015 %>%
  slice(1:(n()-8)) %>%
  stretch_tsibble(.init = 3, .step = 1)

fc <- google_2015_tr %>%
  model(RW(Close ~ drift())) %>%
  forecast(h = 8) %>%
  group_by(.id) %>%
  mutate(h = row_number()) %>%
  ungroup()

fc %>%
  accuracy(google_2015, by = c("h", ".model")) %>%
  ggplot(aes(x = h, y = RMSE)) +
  geom_point()
```

### dev_data single region

```{r}
tictoc::tic()
dev_data_tr <- dev_data %>% 
  filter(regional_unit == "W92003") %>%
  mutate(datename = yearmonth(datename)) %>% 
  select(datename, n) %>%
  fill_gaps() %>% 
  slice(1:(n()-6)) %>% 
  stretch_tsibble(.init = 36, .step = 1)

plan(multisession)
fit <- dev_data_tr %>%
  fableModels()

fc <- fit %>%
  forecast(h = 6)
plan(sequential)
tictoc::toc()

fc %>%
    accuracy(mutate(filter(dev_data, regional_unit == "W92003"),
                    datename = yearmonth(datename)))

plan(multisession)
bind_rows(
  fc %>%
    accuracy(mutate(filter(dev_data, regional_unit == "W92003"),
                    datename = yearmonth(datename))),
  filter(dev_data, regional_unit == "W92003") %>% 
    mutate(datename = yearmonth(datename)) %>%
    fableModels() %>%
    accuracy())
plan(sequential)

fc %>%
  group_by(.id, .model) %>%
  mutate(h = row_number()) %>%
  ungroup() %>%
  accuracy(mutate(filter(dev_data, regional_unit == "W92003"),
                    datename = yearmonth(datename))) %>%
  ggplot(aes(x = .model, y = RMSE, fill = as.factor(.id))) +
  geom_point() +
  facet_grid(~"regional_unit")

dev_data_tr %>% 
  filter(.id == 1) %>% 
  .$datename %>% 
  unique()

fc %>%
  group_by(.id, .model) %>%
  ungroup() %>% 
    accuracy(mutate(filter(dev_data, regional_unit == "W92003"),
                    datename = yearmonth(datename)))
ggplot(aes(x = datename, y = ))

```

* 1 practice, 6 months, 15 windows = 55 secs
* With 1 practice, init = 36, step = 1 you end up with 360 obs of 3 variables (datename, n, .id) in the training data set
* with 1 practice, init = 35, step = 1 you end up with 395 obs of 3 variables (same as above) in the training data set

### multiple regions, single window

```{r}
tictoc::tic()
dev_data_tr <- dev_data %>% 
  mutate(datename = yearmonth(datename)) %>% 
  select(datename, n, regional_unit) %>%
  group_by(regional_unit) %>% 
  fill_gaps() %>% 
  slice(1:(n()-6), .preserve = TRUE) %>% # group_by(regional_unit) and .preserve must be true to maintain structure
  stretch_tsibble(.init = 35, .step = 1)

regions <- unique(dev_data_tr$regional_unit)
.ids <- unique(dev_data_tr$.id)

plan(multisession)
fable_fit <-
  future_map_dfr(.x = regions,
                  .f = ~regionalFableFits(filter(dev_data_tr, .id == 1), .x),
                  .options = furrr_options(seed = TRUE))

message("fable fits complete")

fc <- forecast(fable_fit, h = 6)

message("fabel forecasts complete")

plan(sequential)
tictoc::toc()

fc %>% 
    accuracy(mutate(dev_data,
                    datename = yearmonth(datename)), by = c("regional_unit", ".model"))

bind_rows(
  fc %>%
    accuracy(mutate(dev_data, datename = yearmonth(datename))),
  dev_data %>%
    mutate(datename = yearmonth(datename)) %>%
    fableModels() %>%
    accuracy())

```

* With 2 practices, you end up with 1005 obs of 4 variables (datename, n, regional_unit, .id)
* 100 practices, 6 months, 1 window = 285 seconds

```{r}
dev_data_tr_36 <- dev_data %>% 
  mutate(datename = yearmonth(datename)) %>% 
  select(datename, n, regional_unit) %>%
  group_by(regional_unit) %>% 
  fill_gaps() %>% 
  slice(1:(n()-6), .preserve = TRUE) %>% # group_by(regional_unit) and .preserve must be true to maintain structure
  stretch_tsibble(.init = 36, .step = 1)

dev_data_tr_36 %>% 
  as_tibble() %>% 
  group_by(.id, regional_unit) %>% 
  count

dev_data_tr_35 <- dev_data %>% 
  mutate(datename = yearmonth(datename)) %>% 
  select(datename, n, regional_unit) %>%
  group_by(regional_unit) %>% 
  fill_gaps() %>% 
  slice(1:(n()-6), .preserve = TRUE) %>% # group_by(regional_unit) and .preserve must be true to maintain structure
  stretch_tsibble(.init = 35, .step = 1)

dev_data_tr_35 %>% 
  as_tibble() %>% 
  group_by(.id, regional_unit) %>% 
  count
```

### dev_data

```{r, loo-cv}
#  .init = nrow(df)-cv_dist and slice 1:n()-cv_dist when .step = 1 allows 
  # parallelFableFits() for .ids[1:length(ids)-cv_dist]

parallelFableRegionalFits <- function(data, regions, window){

  print(paste("fable regional fit", window))
  future_map_dfr(.x = regions,
                 .f = ~regionalFableFits(filter(data, .id == window), .x),
                 .options = furrr_options(seed = TRUE))
}

trainingFableRegionalFits <- function(data, cv_dist = 6, init = 36, step = 1){
  # This function will make use of parallel processing if set up using plan()
  # Set variables
  cv_dist <- cv_dist

  # Create training data
  data_tr <- data %>% 
    # group_by(regional_unit) and slice(..., .preserve = TRUE) to maintain structure
    mutate(datename = yearmonth(datename)) %>% 
    select(datename, n, regional_unit) %>%
    group_by(regional_unit) %>% 
    fill_gaps() %>% 
    slice(1:(n()-cv_dist), .preserve = TRUE) %>% 
    stretch_tsibble(.init = init, .step = step)

  # Set helper variables
  regions_ids <- unique(data_tr$regional_unit)
  ids <- unique(data_tr$.id)

  # Call functions
  ## Fits
  fable_fit <- 
    map(.x = ids,#(length(ids)-cv_dist)],
            .f = ~parallelFableRegionalFits(dev_data_tr, 
                                            regions = regions_ids, window = .x),
            .id = "window")
  message("fable fits complete")

  ## Forecasts
  fc <- forecast(fable_fit, h = cv_dist)
  message("fable forecasts complete")

  # Return
  res <- list(fits = fable_fit, forecasts = fc)
  return(res)
}

# # As script
# tictoc::tic()
# cv_dist <- 6
# dev_data_tr <- dev_data %>% # Training data for CV
#   mutate(datename = yearmonth(datename)) %>% 
#   select(datename, n) %>%
#   fill_gaps() %>% 
#   slice(1:(n()-cv_dist)) %>% 
#   stretch_tsibble(.init = 36, .step = 1)
# 
# regions_ids <- unique(dev_data_tr$regional_unit)
# ids <- unique(dev_data_tr$.id)
# 
# plan(multisession)
# 
# 
# fable_fit <- 
#   map_dfr(.x = ids[1:(length(ids)-cv_dist)],
#           .f = ~parallelFableRegionalFits(dev_data_tr, 
#                                           regions = regions_ids, window = .x),
#           .id = "window")
# 
# message("fable fits complete")
# 
# fc <- forecast(fable_fit, h = cv_dist)
# 
# message("fabel forecasts complete")
# 
# plan(sequential)
# tictoc::toc()

# As function with default values
plan(multisession)
system.time(fable_cv <- trainingFableRegionalFits(data = dev_data))
plan(sequential)

# As function with different init value
plan(multisession)
system.time(fable_cv_2 <- trainingFableRegionalFits(data = dev_data, init = 37)) 
system.time(fable_cv_2 <- trainingFableRegionalFits(data = dev_data, init = 35))
# Fails with init < 36
plan(sequential)

fable_cv_2$forecasts %>% 
  accuracy(mutate(dev_data,
                    datename = yearmonth(datename)), by = c("regional_unit", ".model"))

fc %>% 
    accuracy(mutate(dev_data,
                    datename = yearmonth(datename)), by = c("regional_unit", ".model"))

bind_rows(
fc %>% 
    accuracy(mutate(dev_data,
                    datename = yearmonth(datename)), by = c("regional_unit", ".model")),
          dev_data %>%
            mutate(datename = yearmonth(datename)) %>%
            fableModels() %>%
            accuracy()) %>% 
  group_by(.model, .type) %>% 
  summarise(across(where(is.numeric), mean))

```

* ~108 seconds to CV 2 practices for 6 months across 9 windows
* ~270 seconds to CV 10 practices for 6 months across 9 windows
  * If this extrapolates in a  linear fashion, 400 practices would take `r 40*270/60/60` hours to calculate
  * Based on the 2 practice timing, this seems unlikely and the final time should be less than that
* Why is id 9 the maximum that will work?

### fitting and cv in the same function

```{r}
tictoc::tic()
regions <- unique(dev_data$regional_unit)
cv_dist <- 6
init <- 36
step <- 1
data <- 
  dev_data %>% 
  filter(regional_unit == regions[1]) %>% 
  select(regional_unit, datename, n) %>% 
  mutate(datename = yearmonth(datename)) %>% 
  fill_gaps() %>% 
  slice(1:(n()-cv_dist), .preserve = TRUE) %>% 
  stretch_tsibble(.init = init, .step = step)

fits <- fableModels(data)

fc <- fits %>% forecast(h = cv_dist)

acc <- fc %>% 
  accuracy(mutate(filter(dev_data, regional_unit == regions[[1]]),
                    datename = yearmonth(datename))) %>% 
  arrange(RMSE) %>% 
  slice(1)
acc

tictoc::toc()
```

* `r mean(c(87.7, 76.6, 77.3, 82.517))`

```{r}

fableModels(dev_data)
regionalFableFits(dev_data, Arfon)

system.time(best_models_1 <- regionalCV(dev_data))
system.time(best_models_2 <- regionalCV(dev_data, init = 30))
system.time(best_models_3 <- regionalCV(dev_data, step = 2))

system.time(best_oc_models1 <- regionalCV(oc_monthly_quantity_gp_pre_cv19))

best_oc_models1 %>% 
  saveRDS(file = "../data/oc_gp_cv_models.rds")

best_models_1 %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = RMSE)) +
  geom_histogram()

best_models_1 %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))
```

* 1 regional_unit
  * Mean time `r mean(c(78.5, 77.7))` s w/o prophet
  * Mean time `r mean(c(142.466, 147.631))` s w prophet
* 2 regional units
  * min = 671.0648, lq = 691.4213, mean = 699.9314, median = 700.2545, uq = 710.868, max = 726.0485, 
* 100 regional units
  * Mean time `r mean(c(66, 61))` mins
* 400 regional units
  * `r lubridate::seconds_to_period(29663.27)` hours

### prophet::cross_validation()

```{r}
cross_validation(select(oc_gp_cluster_fcts$fits, c(regional_unit, starts_with("oc_proph"))), horizon = , units = "years", initial = 3)
```

## Model selection

```{r}
## accuracy of each model on each regional_unit
accuracies <- joint_fits$fits %>% 
  accuracy()

bestModelFit <- function(data, measure){
  measure = enquo(measure)

  data <- janitor::clean_names(data)

  data %>% 
    group_by(regional_unit) %>% 
    summarise("{{ measure }}" := min(!!measure), .groups = "drop") %>% 
    select(regional_unit, !!measure) %>% 
    left_join(data, by = c("regional_unit", as_name(measure))) %>% # doesn't provide the single match if models are tied 
    select(regional_unit, model, everything(), -type)

}

bestModelFit(accuracies, rmse)
```

## Joint Fable and Prophet fits & forecast

* Take a tsibble and fit both the fable and prophet models to return a single dataframe of models that can be used for forecasting

```{r}
library(progressr)
handlers("progress")

joint_fits <- regionalFitsJoint(data = dev_data)

joint_fits
```

## Close multisession workers
```{r}
plan(sequential)
```

## Other

```{r, eval=FALSE}
x <- dev_data %>% 
  mutate(datename = yearmonth(datename)) %>% 
  as_tsibble(index = datename) %>% 
  model(oc_prophet_model1, oc_prophet_model2, oc_prophet_model3, 
        oc_prophet_model4, oc_prophet_model5)

x2 <- forecast(x, h = "5 months")

mean(quantile(x = x2$n, p = c(.1)))
mean(quantile(x = x2$n, p = c(.5)))
mean(quantile(x = x2$n, p = c(.9)))

x3 <- 
  x2 %>% 
  mutate(LB = quantile(n, p = .05),
         .median = quantile(n, p = .5),
         UB = quantile(n, p = .95))

x3 %>% 
  mutate(model = as.factor(.model)) %>% 
  filter(regional_unit == "W91027") %>% 
  ggplot(aes(x = datename, y = .median, colour = as.factor(model))) + 
  geom_line(aes(colour = as.factor(model)), lty = 2) +
  geom_ribbon(aes(ymin = LB, ymax = UB, fill = as.factor(model)), alpha = .3) +
  geom_line(aes(y = n, x = datename), data = filter(oc_monthly_quantity_gp %>% mutate(datename = yearmonth(datename)), regional_unit == "W91027"))

###

proph_fcst1 <-
  proph_fit %>% 
  forecast(h = "5 months")
autoplot(filter(proph_fcst1, regional_unit == "W91027"))

proph_fcst1 %>% 
  ggplot(aes(x = datename, y = .mean, colour = as.factor(regional_unit))) +
  geom_line() +
  facet_wrap(~as.factor(regional_unit))

filter(proph_fcst1, regional_unit == "W92003") %>% 
  autoplot(filter(oc_monthly_quantity_gp, regional_unit == "W92003"))

###

df <- oc_monthly_quantity_gp_pre_cv19 %>% 
  transmute(ds = yearmonth(datename), y = n)
m <- prophet::prophet(df, seasonality.mode = "additive")
future <- make_future_dataframe(m, periods = 6, freq = 'month')
fcst <- predict(m, future)
prophet_plot_components(m, fcst)

oc_monthly_quantity_gp %>% 
  ggplot(aes(x = datename, y = n)) +
  geom_line() +
  geom_ribbon(aes(ymin = c(rep(NA,50), fcst[51:55,]$yhat_lower),
                  ymax = c(rep(NA,50), fcst[51:55,]$yhat_upper)),
              alpha = .1, fill = "blue") +
  geom_line(aes(y = c(fcst[1:50,]$yhat, rep(NA, 5))),
            col = "blue", lty = 2) +
  geom_line(aes(y = c(rep(NA,50), fcst[51:55,]$yhat)),
            col = "blue", lty = 3) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

## parallel joint forecasts

Why does only one part of this run in parallel?

```{r}
regionalJointFcsts <- function(data, h ="5 months"){
  regions <- unique(data$regional_unit)
  plan(multisession, gc = TRUE)
  fable_fit <-
    future_map_dfr(.x = regions,
                   .f = ~fableModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("fable fits complete")

  fable_fcst <- 
    future_map_dfr(.x = regions,
                   .f = ~as_tibble(forecast(filter(fable_fit, 
                                                   regional_unit ==.x),
                                  h = h)),
                   .options = furrr_options(seed = TRUE))
    # fable_fit %>%
    # forecast(h = h) %>%
    # as_tibble()
  message("fable forecasts complete")

  prophet_fit <-
    future_map_dfr(.x = regions,
                   .f = ~prophetModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("prophet fits complete")

  prophet_fcst <-
        future_map_dfr(.x = regions,
                   .f = ~as_tibble(forecast(filter(prophet_fit, 
                                                   regional_unit ==.x),
                                  h = h)),
                   .options = furrr_options(seed = TRUE))
    # forecast(prophet_fit, h = h) %>%
    # as_tibble()
  # Some prophet forecasts are giving daily forecasts. We want monthly forecasts.
  # This may also be influencing accuracy

  message("prophet forecasts complete")
  plan(sequential)
  # prophetModels now being run in parallel. Needed to reinstall prophet from
  # source

  joint_fcst <-
    bind_rows(fable_fcst, prophet_fcst) %>%
    mutate(lb_95 = quantile(n, .025),
           lb_80 = quantile(n, .1),
           med = median(n),
           ub_80 = quantile(n, .9),
           ub_95 = quantile(n, .975),
           sd = distributional::variance(n)^2,
           dist = n,
           n = NULL)

  return(joint_fcst)
}

system.time(regionalJointFcsts(data = dev_data))
```

84.619 seconds when opening and closing plan(multisession) over 5 regional units vs 59.108 when leaving them open