### 2021-03-04 r_notebooks/gp_dimensionality.Rmd
Exploring the dimensionality of GP data

---
title: "gp_dimensionality"
author: "Will A. S. Hardy"
date: "26/02/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lavaan)

set.seed(376)
knitr::opts_chunk$set(echo = TRUE)
gp <- 
  read_rds("../OneDrive/2_interimdata/8_linkeddata/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL,
         practice_id = as_factor(practiceid),
         dispensing = as_factor(dispensing),
         hboard = as_factor(str_remove_all(hboard, "W110000")),
         cluster = as_factor(cluster)) %>% 
  ungroup() %>% 
  filter(year == 2020) %>% 
  select(-c(practiceid, postcode, easting, northing, lsoacode, year)) %>% 
  janitor::clean_names() %>%
  mutate(across(chd:flu, .fns = ~.x*100))

gp1 <- 
  slice_sample(gp, prop = .5) 
gp2 <- 
  filter(gp, practice_id %in% gp1$practice_id)
```

```{r, inspect-data}
gp %>%
  select(chd:flu) %>% 
  psych::describe() %>% 
  arrange(sd)

gp %>%
  select(chd:flu) %>% 
  DataExplorer::plot_histogram()
```

## QOF disease register EFA

```{r}
df <- select(gp1, chd:flu)
psych::fa.parallel(df)

three_fac <- psych::fa(df, nfactors = 3, rotate = "varimax")
four_fac <- psych::fa(df, nfactors = 4, rotate = "varimax")
five_fac <- psych::fa(df, nfactors = 5, rotate = "varimax")
print(three_fac$loadings, cutoff = .2)
print(four_fac$loadings, cutoff = .2)
```

```{r}

summary(three_fac)
summary(four_fac)
summary(five_fac)
```

## QOF disease register CFA

```{r, cfa-1}
cfa1 <- 
'
  level: 1
  cardio      =~ af + chd + hf + hyp + str
  pulmonary   =~ flu + ast + copd
  cognitive   =~ dem +lrn + men
  f4          =~ diab + obi
  age         =~ can + ost + pal

  level: 2 # Saturated correlation matrix
  # af    ~~ af + chd + hf + hyp + str + flu + ast + copd
  # chd   ~~ chd + hf + hyp + str + flu + ast + copd
  # hf    ~~ hf + hyp + str + flu + ast + copd
  # hyp   ~~ hyp + str + flu + ast + copd
  # str   ~~ str + flu + ast + copd
  # flu   ~~ flu + ast + copd
  # ast   ~~ ast + copd
  # copd  ~~ copd

  af    ~~ af + chd + hf + hyp + str + flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  chd   ~~ chd + hf + hyp + str + flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  hf    ~~ hf + hyp + str + flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  hyp   ~~ hyp + str + flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  str   ~~ str + flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  flu   ~~ flu + ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  ast   ~~ ast + copd + lrn + men + diab + obi + can + dem + ost + pal
  copd  ~~ copd + lrn + men + diab + obi + can + dem + ost + pal
  lrn   ~~ lrn + men + diab + obi + can + dem + ost + pal
  men   ~~ men + diab + obi + can + dem + ost + pal
  diab  ~~ diab + obi + can + dem + ost + pal
  obi   ~~ obi + can + dem + ost + pal
  can   ~~ can + dem + ost + pal
  dem   ~~ dem + ost + pal
  ost   ~~ ost + pal
  pal   ~~ pal
'
# rar epi

fit1 <- sem(model = cfa1, data = gp1, cluster = "cluster", estimator = "MLR")

summary(fit1, fit.measures=TRUE, standardized = TRUE)

lavInspect(fit1, "icc")
```