### 2021-06-01 writing/paper.Rmd

---
title             : "Variation in Primary Care Prescribing in Wales due to COVID-19"
shorttitle        : "Welsh Primary Care COVID Prescribing"

author: 
  - name          : "Will Hardy"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Susan Murphy"
    affiliation   : "3"
  - name          : "Dyfrig Hughes"
    affiliation   : "1"
  - name          : "Daniel Hill-McManus"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Centre for Health Economics & Medicines Evaluation, Bangor University, LL55 2PZ"
  - id            : "2"
    institution   : "Institute for the Psychology of Elite Performance, Bangor University, LL55 2PZ"
  - id            : "3"
    institution   : "Betsi Cadwaladr University Health Board"

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib", "../references/library.bib"]

header-includes:
   - \usepackage{caption}
csl               : "../references/apa.csl"
always_allow_html : true
output            : 
  - html_document

---

```{r setup, include = FALSE}
## Packages
# devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(fasster)
library(feasts)
library(furrr)
library(rebus)
library(serCymruTools)
library(knitr)
library(kableExtra)
library(geepack)
library(papaja)

r_refs("r-references.bib")

## Data
gp <- 
  read_rds("../data/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL,
         practice_id = as.character(practiceid)) %>% 
  ungroup() %>% 
  janitor::clean_names() %>% 
  select(practice_id, year, npat)

# CV model accuracies
cv_model_files <- 
  paste0("../data/gp_cv_models/", 
         list.files("../data/gp_cv_models/"))
names(cv_model_files) <- 
  cv_model_files %>% 
  str_remove_all("../data/gp_cv_models/") %>% 
  str_remove_all("_gp_cv_models.rds")

# Forecast - observed
differences_files <- 
  list.files("../data/differences/")
names(differences_files) <- 
  str_remove_all(differences_files, "_discreps.rds")

differences <- 
  map_dfr(.x = differences_files,
          .f = ~read_rds(paste0("../data/differences/", .x)),
          .id = "drug") %>% 
  mutate(drug = case_when(drug == "contra" ~ "oral",
                          !drug == "contra" ~ drug),
         drug = relevel(as_factor(drug), ref = "all_drugs"),
         sd = sd^.25) # Error in serCymruTools when calculated

# GP extract data
data_files <-
  paste0("../data/prescribing_data/", list.files("../data/prescribing_data")) 
names(data_files) <- 
  str_remove_all(list.files("../data/prescribing_data"), ".rds")

gp_extract <- 
  map_dfr(.x = data_files,
          .f = ~(read_rds(.x) %>% 
            filter(practice_id %in% unique(differences$regional_unit))),
          .id = "drug") 
```

```{r, analysis-preferences}
# Seed for random number generation
set.seed(32)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
kable_opts <- options(knitr.kable.NA = "-")
```

## Figure 1

```{r, figure-1, fig.cap="Data processing flow."}
include_graphics("figures/data_flow.pdf")
```

## Table 2

```{r, table-2}
read_rds("../data/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL,
         practice_id = as.character(practiceid)) %>% 
  ungroup() %>% 
  janitor::clean_names() %>%  
  filter(year == 2020) %>% 
  select(npat, ngp, dispensing, wimd2019:flu) %>% 
  mutate(across(p_male:flu, ~.x*100)) %>% 
  psych::describe() %>% 
  select(mean, sd, median, min, max) %>% 
  kable(digits = 2)
```

## Figure 2

```{r, figure-2, fig.cap="Distribution of the difference between observed and forecast prescribing, expressed as a percentage of forecast prescribing. Grey line marks forecast level of prescribing. Pink shading covers duration of Wales' \"stay at home\" order. Please note the difference y-axis scales across the plots.", dpi=450}
differences_quantiles <- 
  differences %>% 
  na.omit() %>% 
  group_by(datename, drug) %>% 
  summarise(mean_prop = mean(prop, na.rm = T),
            med_prop = median(prop, na.rm = T),
            sd_prop = sd(prop, na.rm = T),
            se_prop = sd(prop, na.rm = T)/sqrt(length(prop)),
            enframe(quantile(prop, seq(.1,.9,.2)), 
                    "Percentile", "prop"),
            .groups = "drop") %>% 
  mutate(datename = month(datename),
         prop = 100*(prop - 1)) # Percentage change instead of proportion

differences_quantiles %>%
  mutate(datename = factor(x = datename, levels = 1:8, 
                           labels = month.abb[1:8])) %>% 
  mutate(drug = case_when(drug == "all_drugs" ~ "All",
                          drug == "ace" ~ "ACE inhibitors",
                          drug == "adreno" ~ "Adrenoceptor agonists",
                          drug == "azithro" ~ "Azithromycin",
                          drug == "contra" ~ "Oral contraceptives",
                          drug == "cortico" ~ "Corticosteroids",
                          drug == "hcq" ~ "Hydroxycholoroquine",
                          drug == "nsaid" ~ "NSAIDs",
                          drug == "paracet" ~ "Paracetamol",
                          drug == "ssri" ~ "SSRIs",
                          drug == "vitd" ~ "Vitamin D",
                          drug == "warfarin" ~ "Warfarin"),
         drug = relevel(as_factor(drug), ref = "All")) %>%
  ggplot(aes(x = datename, col = drug, 
             group = Percentile)) +
  geom_rect(aes(xmin = "Mar", xmax = "Jul", ymin = -Inf, ymax = Inf),
            fill = 'pink', col = 0,alpha = 0.03) +
  geom_line(aes(y = prop, lty = Percentile)) +
  geom_hline(aes(yintercept = 0), alpha = .5) +
  facet_wrap(~drug, scales = "free") + 
  guides(size = "legend", colour = "none") + 
  theme(legend.position = "bottom", text = element_text(size = 10)) +
  scale_linetype_manual(breaks = c("10%", "30%", "50%"), # Remove 70% and 90% from legend
                        values = c(3,2,1,2,3), # Replicate line type
                        labels = c("10% lower; 90% upper", # Amalgamate line type labels
                                   "30% lower; 70% uppper", "50%")) +
  xlab("Month") +
  ylab("Percentage difference from forecast prescribing")

  ggsave(filename = "Draft 2/Figures/Figure 2.png", device = "png",
         width = 10.5)

```
```{r}
differences_quantiles %>%
  mutate(datename = factor(x = datename, levels = 1:8, 
                           labels = month.abb[1:8])) %>% 
  mutate(drug = case_when(drug == "all_drugs" ~ "All",
                          drug == "ace" ~ "ACE inhibitors",
                          drug == "adreno" ~ "Adrenoceptor agonists",
                          drug == "azithro" ~ "Azithromycin",
                          drug == "contra" ~ "Oral contraceptives",
                          drug == "cortico" ~ "Corticosteroids",
                          drug == "hcq" ~ "Hydroxycholoroquine",
                          drug == "nsaid" ~ "NSAIDs",
                          drug == "paracet" ~ "Paracetamol",
                          drug == "ssri" ~ "SSRIs",
                          drug == "vitd" ~ "Vitamin D",
                          drug == "warfarin" ~ "Warfarin"),
         drug = relevel(as_factor(drug), ref = "All")) %>% 
  filter(drug == "Adrenoceptor agonists" & 
           datename == "Mar")

```

## Table 3

```{python, composite-sd-function-py, eval=FALSE}

from __future__ import print_function, division  # requires python 2.6 or later (2.7 or later preferred)
import math

__all__ = ['iterable', 'avg', 'sample_SD', 'composite_SD']

def iterable(obj):
    '''True iff obj is iterable: a list, tuple, or string.'''
    return hasattr(obj, '__contains__')

def avg(samples):
    if len(samples) >= 1:
        return sum(samples) / len(samples)
    return float('nan')

def sample_SD(samples):
    '''input is an array of samples; result is the standard deviation'''
    mean = avg(samples)
    sum_of_squared_deviations = 0;
    sd = 0
    if len(samples) >= 2:
        for datum in samples:
            sum_of_squared_deviations += ((datum - mean) * (datum - mean));
        sd = math.sqrt(sum_of_squared_deviations / (len(samples)-1) );
    return sd

def composite_SD(means, SDs, ncounts):
    '''Calculate combined standard deviation via ANOVA (ANalysis Of VAriance)
       See:  http://www.burtonsys.com/climate/composite_standard_deviations.html
       Inputs are:
         means, the array of group means
         SDs, the array of group standard deviations
         ncounts, number of samples in each group (can be scalar
                  if all groups have same number of samples)
       Result is the overall standard deviation.
    '''
    G = len(means)  # number of groups
    if G != len(SDs):
        raise Exception('inconsistent list lengths')
    if not iterable(ncounts):
        ncounts = [ncounts] * G  # convert scalar ncounts to array
    elif G != len(ncounts):
        raise Exception('wrong ncounts list length')

    # calculate total number of samples, N, and grand mean, GM
    N = sum(ncounts)  # total number of samples
    if N <= 1:
        raise Exception("Warning: only " + str(N) + " samples, SD is incalculable")
    GM = 0.0
    for i in range(G):
        GM += means[i] * ncounts[i]
    GM /= N  # grand mean

    # calculate Error Sum of Squares
    ESS = 0.0
    for i in range(G):
        ESS += ((SDs[i])**2) * (ncounts[i] - 1)

    # calculate Total Group Sum of Squares
    TGSS = 0.0
    for i in range(G):
        TGSS += ((means[i]-GM)**2) * ncounts[i]

    # calculate standard deviation as square root of grand variance
    result = math.sqrt((ESS+TGSS)/(N-1))
    return result

samples = range(10)
print(samples)
print('avg=', avg(samples))
sd = sample_SD(samples)
print('sd=', sd)
pt1 = [0,1,2]
pt2 = [3,4]
pt3 = [5,6,7,8,9]
means = [avg(pt1), avg(pt2), avg(pt3)]
SDs = [sample_SD(pt1), sample_SD(pt2), sample_SD(pt3)]
ncounts = [len(pt1), len(pt2), len(pt3)]
sd2 = composite_SD(means, SDs, ncounts)
print('sd2=', sd2)

samples = range(9)
print('avg=', avg(samples))
sd = sample_SD(samples)
print('sd=', sd)
pt1 = [0,1,2]
pt2 = [3,4,5]
pt3 = [6,7,8]
means = [avg(pt1), avg(pt2), avg(pt3)]
SDs = [sample_SD(pt1), sample_SD(pt2), sample_SD(pt3)]
ncounts = 3
sd2 = composite_SD(means, SDs, ncounts)
print('sd2=', sd2)
```

```{r, composite-sd-function-r, echo=FALSE}
avg <- function(samples){
  if (length(samples) >= 1){
    return(sum(samples) / length(samples))
  }
  else{return(NaN)}
}

sample_SD <- function(samples){
  # Input is an vector of samples; result is the standard deviation
  mean = avg(samples)
  sum_of_squared_deviations <-  0
  sd <-  0
  if(length(samples) >= 2){
    for (datum in samples){
      sum_of_squared_deviations <- 
        sum_of_squared_deviations + ((datum - mean) * (datum - mean))
      sd = sqrt(sum_of_squared_deviations / (length(samples)-1))
    }
      return(sd)
  }
}

compositeSD <- function(means, SDs, ncounts){
  #Calculate combined standard deviation via ANOVA (ANalysis Of VAriance)
  # See:  http://www.burtonsys.com/climate/composite_standard_deviations.html
  #    Inputs are:
  #      means, the array of group means
  #      SDs, the array of group standard deviations
  #      ncounts, number of samples in each group (can be scalar
  #               if all groups have same number of samples)
  #    Result is the overall standard deviation.
  G <- length(means) # number of groups
  if(G != length(SDs)) stop("Inconsistent lengths (means & sds")
  if(!is.numeric(ncounts)) {(as.numeric(ncounts))}
  if(G != length(ncounts)) stop("Inconsistent lengths (means & ncounts)")

  # Calculate total number of samples, N, and grand mean, GM
  N = sum(ncounts)  # total number of samples
  if (N <= 1) stop(paste("Warning: only ", N, " samples, SD is incalculable"))
  GM <-  0.0
  for (i in 1:G){
    GM <- GM + means[i] * ncounts[i]
  }
  GM <- GM / N  # grand mean

  # calculate Error Sum of Squares
  ESS <-  0.0
  for (i in 1:G){
    ESS <- ESS + ((SDs[i])**2) * (ncounts[i] - 1)
  }

  # calculate Total Group Sum of Squares
  TGSS <-  0.0
  for (i in 1:G){
    TGSS <- TGSS + ((means[i]-GM)**2) * ncounts[i]
  }

  # calculate standard deviation as square root of grand variance
  result <- sqrt((ESS+TGSS)/(N-1))
  return(result)
}

samples = 0:9
print(paste("avg =", avg(samples)))
sd = sample_SD(samples)
print(paste("sd =", sd))
pt1 = c(0,1,2)
pt2 = c(3,4)
pt3 = c(5,6,7,8,9)
means <- c(avg(pt1), avg(pt2), avg(pt3))
SDs <- c(sample_SD(pt1), sample_SD(pt2), sample_SD(pt3))
ncounts <- c(length(pt1), length(pt2), length(pt3))
sd2 <- compositeSD(means, SDs, ncounts)
print(paste("sd2 =", sd2))

samples = 0:8
print(paste("avg =", avg(samples)))
sd = sample_SD(samples)
print(paste("sd =", sd))
pt1 = c(0,1,2)
pt2 = c(3,4,5)
pt3 = c(6,7,8)
means <- c(avg(pt1), avg(pt2), avg(pt3))
SDs <- c(sample_SD(pt1), sample_SD(pt2), sample_SD(pt3))
ncounts <- c(length(pt1), length(pt2), length(pt3))
sd2 <- compositeSD(means, SDs, ncounts)
print(paste("sd2 =", sd2))

```

```{r, table-3-single-fcts, eval=FALSE}
nat_data <- 
  data_files %>% 
  map_dfr(read_rds, .id = "drug") %>% 
  group_by(drug, datename) %>% 
  summarise(n = sum(quantity), .groups = "drop")

plan(multisession)
precribing_cv_models <- 
  nat_data %>% 
  mutate(datename = yearmonth(dmy(paste("28", datename, sep = "-"))),
         regional_unit = as.factor(drug)) %>% 
  filter(datename < yearmonth(dmy("01-01-2020"))) %>% 
  as_tsibble(index = "datename", key = "regional_unit") %>% 
  serCymruTools::regionalCV(cv_dist = 8, init = 36, step = 3)
plan(sequential)

precribing_cv_models

best_model <- 
  precribing_cv_models %>% 
  group_by(regional_unit) %>% 
  rankModels() %>% 
  slice(1, .preserve = TRUE)

best_mape_over_50 <- 
  best_model %>% 
  filter(MAPE > 50)

plan(cluster)
system.time(prescribing_fcsts <-
                nat_data %>% 
  mutate(datename = yearmonth(dmy(paste("28", datename, sep = "-"))),
         regional_unit = as.factor(drug)) %>% 
  filter(datename < yearmonth(dmy("01-01-2020"))) %>% 
  as_tsibble(index = "datename", key = "regional_unit") %>%
              filter(!regional_unit %in% best_mape_over_50$regional_unit) %>%
              regionalJointFcsts())  
plan(sequential)

best_model_discrep <- 
  best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(prescribing_fcsts,
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(mutate(select(                nat_data %>% 
  mutate(datename = yearmonth(dmy(paste("28", datename, sep = "-"))),
         regional_unit = as.factor(drug)) , 
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - med, # observed - forecast
         prop = n / med) %>% 
  ungroup()

best_model_discrep %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  group_by(regional_unit) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(1, length(sd))),
            .mean = sum(.mean)) %>% 
  mutate(lb = .mean - 1.96*sd,
         ub = .mean + 1.96*sd,
         lb_mean_pct_change = (100 * (n - ub)/ ub),
         mean_pct_change = (100 * (n - .mean)/ .mean),
         ub_mean_pct_change = (100 * (n - lb)/ lb)) %>% 
  select(regional_unit, lb_mean_pct_change, mean_pct_change, ub_mean_pct_change) %>% 
  mutate(pct_change = paste0(printnum(mean_pct_change), 
                             "[", printnum(lb_mean_pct_change), ",", 
                             printnum(ub_mean_pct_change), "]")) %>% 
  select(regional_unit, pct_change)

best_model_discrep %>% 
  mutate(prop = 100*(prop-1)) %>% 
  pivot_wider(id_cols = regional_unit, names_from = datename, values_from = prop) %>% 
  kable(digits = 2)
```

```{r, table-3}
monthly_diffs <- 
  differences %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug, datename) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(nrow(.),length(.mean))),
            .mean = sum(.mean),
            .groups = "drop") %>% 
  mutate(lb = .mean - 1.96*sd,
         ub = .mean + 1.96*sd,
         lb_mean_pct_change = (100 * (n - ub)/ ub),
         mean_pct_change = (100 * (n - .mean)/ .mean),
         ub_mean_pct_change = (100 * (n - lb)/ lb)) %>% 
  mutate(datename = month(datename),
         pct_change = paste0(printnum(mean_pct_change), 
                             "[", printnum(lb_mean_pct_change), ",", 
                             printnum(ub_mean_pct_change), "]")) %>% 
  pivot_wider(id_cols = drug, names_from = datename, values_from = pct_change)

lockdown_diffs <- 
  differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(1,length(.mean))),
            .mean = sum(.mean),
            .groups = "drop") %>% 
  mutate(lb = .mean - 1.96*sd,
         ub = .mean + 1.96*sd,
         lb_mean_pct_change = (100 * (n - ub)/ ub),
         mean_pct_change = (100 * (n - .mean)/ .mean),
         ub_mean_pct_change = (100 * (n - lb)/ lb)) %>% 
  mutate(pct_change = paste0(printnum(mean_pct_change), 
                             "[", printnum(lb_mean_pct_change), ",", 
                             printnum(ub_mean_pct_change), "]")) %>% 
  select(drug, pct_change)

left_join(lockdown_diffs, monthly_diffs) %>%
  kable(col.names = c("Drug", "Lockdown", month.abb[1:8]),
        align = "lrrrrrrrrr",
        booktabs = TRUE,
        caption = "Monthly percentage change in prescribing") %>%
  add_header_above(header = c(" " = 1,
                              "Mean percentage change [95% CI]" = 9)) %>%
  landscape()
```

```{r, eval=FALSE}
differences %>% 
  filter(drug == "adreno" &
           datename == yearmonth("2020 Mar")) %>% 
  ggplot(aes(x = prop-1)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(prop-1)))

differences %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug, datename) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(nrow(.),length(.mean))),
            .mean = sum(.mean),
            .groups = "drop") %>% 
  filter(drug == "adreno" &
           datename == yearmonth("2020 Mar"))

differences %>% 
  filter(drug == "adreno" &
           datename == yearmonth("2020 Mar")) %>% 
  summarise(.mean = sum(.mean),
            n = sum(n),
            prop = mean(prop))

differences %>% 
  filter(drug == "adreno" &
           datename == yearmonth("2020 Mar")) %>% 
  mutate(prop_correct = if_else(prop == n/med, TRUE, FALSE))
```

```{r, eval=FALSE}
differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  group_by(drug) %>% 
  summarise(n = sum(n),
            .mean = sum(.mean)) %>% 
  mutate(pct_change = 100 * (n - .mean ) / .mean)
```

```{r, eval=FALSE}
differences %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug) %>% 
  summarise(across(where(is.numeric),.fns = max))
```

```{r, eval=FALSE}
gp_extract %>% 
  filter(str_detect(datename, "-19")) %>% 
  group_by(drug) %>% 
  summarise(n = sum(quantity)/1e6) %>% 
  left_join(lockdown_diffs %>% 
  select(drug, pct_change), by = "drug")
```

```{r, eval=FALSE}
differences %>% 
  group_by(drug) %>% 
  summarise(discrepancy = sum(discrepancy, na.rm = TRUE),
            n = sum(n, na.rm = TRUE),
            pct = 100*discrepancy / n)

differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  group_by(drug) %>% 
  summarise(discrepancy = sum(discrepancy, na.rm = TRUE),
            n = sum(n, na.rm = TRUE),
            pct = 100*discrepancy / n) %>% 
  ggplot(aes(x = reorder(drug, pct), y = pct)) +
  geom_col()
```