### 2021-09-09 writing/paper.Rmd

---
title: "Variation in Primary Care Prescribing in Wales due to COVID-19"
author:
- name: Will Hardy
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: my@email.com
- name: Susan Murphy
  affiliation: '3'
- name: Dyfrig Hughes
  affiliation: '1'
- name: Daniel Hill-McManus
  affiliation: '1'
shorttitle: Welsh Primary Care COVID Prescribing
output:
  html_document: default
  pdf_document: default
keywords: keywords
wordcount: X
bibliography:
- r-references.bib
- ../references/library.bib
header-includes: \usepackage{caption}
csl: ../references/apa.csl
always_allow_html: yes
affiliation:
- id: '1'
  institution: Centre for Health Economics & Medicines Evaluation, Bangor University,
    LL55 2PZ
- id: '2'
  institution: Betsi Cadwaladr University Health Board
---

```{r setup, include = FALSE}
## Packages
# devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(fable)
library(feasts)
library(furrr)
library(rebus)
library(serCymruTools)
library(knitr)
library(kableExtra)
library(geepack)
library(papaja)
library(fuzzyjoin)

r_refs("r-references.bib")

plan(cluster)
## Data
gp <- 
  read_rds("../data/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL,
         practice_id = as.character(practiceid)) %>% 
  ungroup() %>% 
  janitor::clean_names() %>% 
  select(practice_id, year, npat)

# CV model accuracies
cv_model_files <- 
  paste0("../data/gp_cv_models/", 
         list.files("../data/gp_cv_models/"))
names(cv_model_files) <- 
  cv_model_files %>% 
  str_remove_all("../data/gp_cv_models/") %>% 
  str_remove_all("_gp_cv_models.rds")

# Forecast - observed
differences_files <- 
  list.files("../data/differences/")
names(differences_files) <- 
  str_remove_all(differences_files, "_discreps.rds")

differences <- 
  furrr::future_map_dfr(.x = differences_files,
          .f = ~read_rds(paste0("../data/differences/", .x)),
          .id = "drug") %>% 
    mutate(.mean = mean(dist),
           drug = relevel(as_factor(drug), ref = "quant_all_drugs"))
plan(sequential)

# GP extract data
data_files <-
  paste0("../data/prescribing_data/", list.files("../data/prescribing_data"))
names(data_files) <-
  str_remove_all(list.files("../data/prescribing_data"), ".rds")
# 
# gp_extract <- 
#   map_dfr(.x = data_files[1:11],
#           .f = ~(read_rds(.x)),
#           .id = "drug")
```

```{r, analysis-preferences}
# Seed for random number generation
set.seed(32)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = FALSE, echo = FALSE)
kable_opts <- options(knitr.kable.NA = "-")
```

## Figure 1

```{r, figure-1, fig.cap="Data processing flow."}
include_graphics("figures/data_flow.pdf")
```

## Table 2

```{r, table-2}
tibble(regional_unit = unique(differences$regional_unit)) %>% 
  left_join(read_rds("../data/gp_agesex_qof_combined.rds") %>% 
              mutate(npat = N,
                     N = NULL,
                     regional_unit = as.character(practiceid)) %>% 
              ungroup() %>% 
              janitor::clean_names() %>%  
              filter(year == 2020)) %>% 
  select(npat, ngp, dispensing, wimd2019:flu) %>% 
  mutate(across(p_male:flu, ~.x*100)) %>% 
  tibble() %>% 
  psych::describe(fast = TRUE) %>% 
  kable(digits = 2, caption = "Summary of prescribing data and GP practice characteristics")
```

```{r}
differences %>% 
  dplyr::group_by(drug) %>% 
  dplyr::count() %>% 
  mutate(n = n/8)

monthlyPrescribing <- 
  function(data){
    data %>% 
      janitor::clean_names() %>% 
      mutate(datename = dmy(paste0("28-", datename)),
             practice_id = as.character(practice_id)) %>% 
      countRegional(practice_id, datename, quantity) %>% 
      mutate(year = year(datename)) %>%
      as_tibble() %>% 
      inner_join(gp %>% 
                   select(practice_id, year, npat), 
                 by = c("regional_unit" = "practice_id", "year")) %>% 
      mutate(n = 1000*n/npat,
             datename = yearmonth(datename),
             regional_unit = as_factor(regional_unit)) %>% 
      select(-c(npat, year)) %>% 
      as_tsibble(index = datename, key = regional_unit) %>% 
      fill_gaps()
  }

plan(cluster)
pres_2019 <- 
  data_files[c(1:10)] %>% 
  furrr::future_map(read_rds) %>% 
  furrr::future_map(monthlyPrescribing) %>% 
  furrr::future_map_dfr(as_tibble, .id = "drug") %>% 
  filter(year(datename) == 2019)
plan(sequential)

pres_2019 %>% 
  right_join(differences %>% 
               group_by(regional_unit, drug) %>% 
               count() %>% 
               select(-n)) %>% 
  filter(drug != "warfarin_ddd") %>% 
  bind_rows(read_rds(data_files[12]) %>% 
              mutate(datename = yearmonth(dmy(paste0("28-", datename)))) %>% 
              rename(regional_unit = practice_id,
                     n = quantity) %>% 
              filter(year(datename) == 2019) %>% 
              right_join(differences %>% 
                           filter(drug == "warfarin_ddd") %>% 
                           group_by(regional_unit, drug) %>% 
                           count() %>% 
                           select(-n))) %>% 
  group_by(drug) %>% 
  summarise(mean = mean(n, na.rm = TRUE),
            sd = sd(n, na.rm = TRUE),
            min = min(n, na.rm = TRUE),
            max = max(n, na.rm = TRUE),
            n = length(unique(regional_unit))) %>% 
  select(drug, n, everything()) %>% 
  kable(digits = 2)

warf_2019 <- 
  read_rds(data_files[12]) %>% 
  mutate(datename = yearmonth(dmy(paste0("28-", datename)))) %>% 
  rename(regional_unit = practice_id) %>% 
  filter(year(datename) == 2019) %>% 
  right_join(differences %>% 
               filter(drug == "warfarin_ddd") %>% 
               group_by(regional_unit, drug) %>% 
               count() %>% 
               select(-n)) %>%

```

## Figure 2

```{r, figure-2, fig.cap="Distribution of the difference between observed and forecast prescribing, expressed as a percentage of forecast prescribing. Grey line marks forecast level of prescribing. Pink shading covers duration of Wales' \"stay at home\" order. Please note the difference y-axis scales across the plots.", dpi=450}
differences_quantiles <- 
  differences %>% 
  na.omit() %>% 
  dplyr::group_by(datename, drug) %>% 
  dplyr::summarise(mean_prop = mean(prop, na.rm = T),
            med_prop = median(prop, na.rm = T),
            sd_prop = sd(prop, na.rm = T),
            se_prop = sd(prop, na.rm = T)/sqrt(length(prop)),
            enframe(stats::quantile(prop, seq(.1,.9,.2)), 
                    "Percentile", "prop"),
            .groups = "drop") %>% 
  dplyr::mutate(datename = month(datename),
         prop = 100*(prop - 1)) # Percentage change instead of proportion

differences_quantiles %>%
  dplyr::mutate(datename = factor(x = datename, levels = 1:8, 
                           labels = month.abb[1:8])) %>% 
  dplyr::mutate(drug = case_when(drug == "quant_all_drugs" ~ "All",
                          drug == "ddd_ace" ~ "ACE inhibitors",
                          drug == "ddd_adreno" ~ "Beta2 adrenoceptor agonists",
                          drug == "ddd_cortico" ~ "Corticosteroids",
                          drug == "ddd_hcq" ~ "Hydroxycholoroquine",
                          drug == "ddd_nsaid" ~ "NSAIDs",
                          drug == "ddd_oral_contra" ~ "Oral contraceptives",
                          drug == "ddd_paracet" ~ "Paracetamol",
                          drug == "ddd_ssri" ~ "SSRIs",
                          drug == "ddd_vitd" ~ "Vitamin D",
                          drug == "warfarin" ~ "Warfarin",
                          drug == "warfarin_ddd" ~"Warfarin"),
         drug = relevel(as_factor(drug), ref = "All")
         ) %>%
  ggplot(aes(x = datename, col = drug, 
             group = Percentile)) +
  geom_rect(aes(xmin = "Mar", xmax = "Jul", ymin = -Inf, ymax = Inf),
            fill = 'pink', col = 0,alpha = 0.03) +
  geom_line(aes(y = prop, lty = Percentile)) +
  geom_hline(aes(yintercept = 0), alpha = .5) +
  facet_wrap(~drug, scales = "free") + 
  guides(size = "legend", colour = "none") + 
  theme(legend.position = "bottom", text = element_text(size = 10)) +
  scale_linetype_manual(values = c(3,2,1,2,3),
                        # labels = c("10% lower; 90% upper", # Amalgamate line type labels
                        #            "30% lower; 70% upper", "50%"),
                        # #breaks = NULL
                        ) +
  # scale_linetype_manual(breaks = c("10%", "30%", "50%"), # Remove 70% and 90% from legend
  #                       values = c(3,2,1,2,3), # Replicate line type
  #                       labels = c("10% lower; 90% upper", # Amalgamate line type labels
  #                                   "30% lower; 70% upper", "50%")
  #                       ) +
  xlab("Month") +
  ylab("Percentage difference from forecast prescribing")

ggsave(filename = "Draft 4/Figures/Figure 2.png", device = "png", width = 10.5)
```

## Table 3

```{r, composite-sd-function-r, eval=FALSE}
avg <- function(samples){
  if (length(samples) >= 1){
    return(sum(samples) / length(samples))
  }
  else{return(NaN)}
}

sample_SD <- function(samples){
  # Input is an vector of samples; result is the standard deviation
  mean = avg(samples)
  sum_of_squared_deviations <-  0
  sd <-  0
  if(length(samples) >= 2){
    for (datum in samples){
      sum_of_squared_deviations <- 
        sum_of_squared_deviations + ((datum - mean) * (datum - mean))
      sd = sqrt(sum_of_squared_deviations / (length(samples)-1))
    }
      return(sd)
  }
}

compositeSD <- function(means, SDs, ncounts){
  #Calculate combined standard deviation via ANOVA (ANalysis Of VAriance)
  # See:  http://www.burtonsys.com/climate/composite_standard_deviations.html
  #    Inputs are:
  #      means, the array of group means
  #      SDs, the array of group standard deviations
  #      ncounts, number of samples in each group (can be scalar
  #               if all groups have same number of samples)
  #    Result is the overall standard deviation.
  G <- length(means) # number of groups
  if(G != length(SDs)) stop("Inconsistent lengths (means & sds")
  if(!is.numeric(ncounts)) {(as.numeric(ncounts))}
  if(G != length(ncounts)) stop("Inconsistent lengths (means & ncounts)")

  # Calculate total number of samples, N, and grand mean, GM
  N = sum(ncounts)  # total number of samples
  if (N <= 1) stop(paste("Warning: only ", N, " samples, SD is incalculable"))
  GM <-  0.0
  for (i in 1:G){
    GM <- GM + means[i] * ncounts[i]
  }
  GM <- GM / N  # grand mean

  # calculate Error Sum of Squares
  ESS <-  0.0
  for (i in 1:G){
    ESS <- ESS + ((SDs[i])**2) * (ncounts[i] - 1)
  }

  # calculate Total Group Sum of Squares
  TGSS <-  0.0
  for (i in 1:G){
    TGSS <- TGSS + ((means[i]-GM)**2) * ncounts[i]
  }

  # calculate standard deviation as square root of grand variance
  result <- sqrt((ESS+TGSS)/(N-1))
  return(result)
}

samples = 0:9
print(paste("avg =", avg(samples)))
sd = sample_SD(samples)
print(paste("sd =", sd))
pt1 = c(0,1,2)
pt2 = c(3,4)
pt3 = c(5,6,7,8,9)
means <- c(avg(pt1), avg(pt2), avg(pt3))
SDs <- c(sample_SD(pt1), sample_SD(pt2), sample_SD(pt3))
ncounts <- c(length(pt1), length(pt2), length(pt3))
sd2 <- compositeSD(means, SDs, ncounts)
print(paste("sd2 =", sd2))

samples = 0:8
print(paste("avg =", avg(samples)))
sd = sample_SD(samples)
print(paste("sd =", sd))
pt1 = c(0,1,2)
pt2 = c(3,4,5)
pt3 = c(6,7,8)
means <- c(avg(pt1), avg(pt2), avg(pt3))
SDs <- c(sample_SD(pt1), sample_SD(pt2), sample_SD(pt3))
ncounts <- c(length(pt1), length(pt2), length(pt3))
sd2 <- compositeSD(means, SDs, ncounts)
print(paste("sd2 =", sd2))

```

```{r, table-3-single-fcts, eval=TRUE}
plan(cluster)
nat_data <- # Data for all regional units combined
  data_files %>% 
  furrr::future_map_dfr(read_rds, .id = "drug") %>% 
  filter(drug != "warfarin_ddd") %>% 
  dplyr::group_by(drug, datename) %>% 
  dplyr::summarise(n = sum(quantity, na.rm = TRUE), .groups = "drop") %>% 
  mutate(datename = yearmonth(dmy(paste("28", datename, sep = "-"))),
         regional_unit = as.factor(drug)) %>% # Calling drug regional_unit to save rewriting all the code
  as_tsibble(index = "datename", key = "regional_unit") %>% 
  fill_gaps()

precribing_cv_models <- 
  nat_data %>% 
  filter(datename < yearmonth(dmy("01-01-2020"))) %>% 
  serCymruTools::regionalCV(cv_dist = 8, init = 36, step = 1)

best_model <- 
  precribing_cv_models %>% 
  group_by(regional_unit) %>% 
  rankModels() %>% 
  slice(1, .preserve = TRUE)

best_mape_over_50 <- 
  best_model %>% 
  filter(MAPE > 50)

#### Monthly forecasts ####
prescribing_fcsts <-
  nat_data %>% 
  filter(datename < yearmonth(dmy("01-01-2020"))) %>% 
  filter(!regional_unit %in% best_mape_over_50$regional_unit) %>%
  regionalJointFcsts()

lockdown_fcsts <- 
  nat_data %>% 
  filter(datename < yearmonth(dmy("01-01-2020"))) %>% 
  filter(!regional_unit %in% best_mape_over_50$regional_unit) %>% 
  fableModels() %>%
  generate(h = 8, times = 1000) %>%
  as_tibble() %>% 
  filter(!datename %in% 
           yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug", "2020 Sept", "2020 Oct"))) %>% 
  group_by(.model, regional_unit, .rep) %>% 
  dplyr::summarise(.sim = sum(.sim)) %>%
  # Store as a distribution
  dplyr::summarise(total = distributional::dist_sample(list(.sim))) %>% 
  mutate(mean = mean(total),
         lb_95 = stats::quantile(total, .025),
         lb_80 = stats::quantile(total, .1),
         med = median(total),
         ub_80 = stats::quantile(total, .9),
         ub_95 = stats::quantile(total, .975),
         sd = distributional::variance(total)^.5,
         total = NULL)

plan(sequential)

#### Monthly discreps ####
best_model_discrep <- 
  best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(mutate(prescribing_fcsts, # Select best forecast
                    datename = yearmonth(datename)), 
             by = c(".model", "regional_unit")) %>%
  left_join(dplyr::mutate(dplyr::select(nat_data, #Join observed dispensing data
                          c(regional_unit, datename, n)), 
                   datename = yearmonth(datename)),
            by = c("datename", "regional_unit")) %>% 
  mutate(discrepancy = n - med, # observed - forecast
         prop = n / med) %>% 
  ungroup()

#### Plot ####
best_model_discrep %>% 
  group_by(regional_unit) %>% 
  dplyr::mutate(across(lb_95:ub_95, ~ 100*(n-.x)/.x)) %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = med)) +
  geom_ribbon(aes(ymin = lb_95, ymax = ub_95), alpha = .3) +
  geom_hline(aes(yintercept = 0), col = "red")+
  facet_wrap(~regional_unit, scales = "free")

#### Table ####
monthly_diffs <- 
  best_model_discrep %>% 
  group_by(regional_unit) %>% 
  dplyr::mutate(across(lb_95:ub_95, ~ 100*(n-.x)/.x)) %>% 
  select(datename, regional_unit, lb_95, med, ub_95) %>% 
  mutate(pct_change = paste0(printnum(med), 
                             "[", printnum(ub_95), ",", 
                             printnum(lb_95), "]")) %>% 
  select(datename, regional_unit, pct_change) %>% 
  pivot_wider(id_cols = regional_unit, names_from = datename, values_from = pct_change)

#### Lockdown discreps ####
lockdown_model_discrep <- 
  best_model %>% 
  select(.model, regional_unit) %>% 
  inner_join(lockdown_fcsts,  # Select best forecast
             by = c(".model", "regional_unit")) %>%
  left_join(nat_data %>% 
              filter(datename > yearmonth("2020 Feb") &
                       datename < yearmonth("2020 Aug")) %>% 
              as_tibble() %>% 
              select(-datename) %>% 
              ungroup() %>% 
              dplyr::group_by(regional_unit) %>% 
              dplyr::summarise(n = sum(n)),
            by = "regional_unit") %>% 
  dplyr::mutate(across(lb_95:ub_95, ~ 100*(n-.x)/.x))

#### Table ####
lockdown_diffs <- 
  lockdown_model_discrep %>% 
  mutate(pct_change = paste0(printnum(med), 
                             "[", printnum(ub_95), ",", 
                             printnum(lb_95), "]")) %>% 
  select(regional_unit, lckdwn = pct_change)

#### Joint table ####
lockdown_diffs %>% 
  left_join(monthly_diffs, by = "regional_unit") %>% 
  kable(col.names = c("Drug", "Lockdown", month.abb[1:8]),
        align = "lrrrrrrrrr",
        booktabs = TRUE,
        caption = "Monthly percentage change in prescribing") %>%
  kable_styling(font_size = 8) %>% 
  add_header_above(header = c(" " = 1,
                              "Mean percentage change [95% CI]" = 9)) %>%
  landscape()
```

```{r, table-3, eval=FALSE}
monthly_diffs <- 
  differences %>%
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug, datename) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(nrow(.),length(.mean))),
            .mean = sum(.mean),
            .groups = "drop") %>% 
  mutate(lb = .mean - 1.96*sd,
         ub = .mean + 1.96*sd,
         lb_mean_pct_change = (100 * (n - ub)/ ub),
         mean_pct_change = (100 * (n - .mean)/ .mean),
         ub_mean_pct_change = (100 * (n - lb)/ lb)) %>% 
  mutate(datename = month(datename),
         pct_change = paste0(printnum(mean_pct_change), 
                             "[", printnum(lb_mean_pct_change), ",", 
                             printnum(ub_mean_pct_change), "]")) %>% 
  pivot_wider(id_cols = drug, names_from = datename, values_from = pct_change)

lockdown_diffs <- 
  differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug) %>% 
  summarise(n = sum(n),
            sd = compositeSD(.mean, sd, rep(1,length(.mean))),
            .mean = sum(.mean),
            .groups = "drop") %>% 
  mutate(lb = .mean - 1.96*sd,
         ub = .mean + 1.96*sd,
         lb_mean_pct_change = (100 * (n - ub)/ ub),
         mean_pct_change = (100 * (n - .mean)/ .mean),
         ub_mean_pct_change = (100 * (n - lb)/ lb)) %>% 
  mutate(pct_change = paste0(printnum(mean_pct_change), 
                             "[", printnum(lb_mean_pct_change), ",", 
                             printnum(ub_mean_pct_change), "]")) %>% 
  select(drug, pct_change)

left_join(lockdown_diffs, monthly_diffs) %>%
  kable(col.names = c("Drug", "Lockdown", month.abb[1:8]),
        align = "lrrrrrrrrr",
        booktabs = TRUE,
        caption = "Monthly percentage change in prescribing") %>%
  kable_styling(font_size = 8) %>% 
  add_header_above(header = c(" " = 1,
                              "Mean percentage change [95% CI]" = 9)) %>%
  landscape()
```

```{r, eval=FALSE}
differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  group_by(drug) %>% 
  summarise(n = sum(n),
            .mean = sum(.mean)) %>% 
  mutate(pct_change = 100 * (n - .mean ) / .mean)
```

```{r, eval=FALSE}
differences %>% 
  select(drug, datename, .mean, sd, n) %>% 
  na.omit() %>% 
  group_by(drug) %>% 
  summarise(across(where(is.numeric),.fns = max))
```

```{r, eval=FALSE}
gp_extract %>% 
  filter(str_detect(datename, "-19")) %>% 
  group_by(drug) %>% 
  summarise(n = sum(quantity, na.rm = TRUE)/1e6) %>% 
  left_join(lockdown_diffs %>% 
  select(drug, pct_change), by = "drug")
```

```{r, eval=FALSE}
differences %>% 
  group_by(drug) %>% 
  summarise(discrepancy = sum(discrepancy, na.rm = TRUE),
            n = sum(n, na.rm = TRUE),
            pct = 100*discrepancy / n)

differences %>% 
  filter(!datename %in% yearmonth(c("2020 Jan", "2020 Feb", "2020 Aug"))) %>% 
  group_by(drug) %>% 
  summarise(discrepancy = sum(discrepancy, na.rm = TRUE),
            n = sum(n, na.rm = TRUE),
            pct = 100*discrepancy / n) %>% 
  ggplot(aes(x = reorder(drug, pct), y = pct)) +
  geom_col()
```