### 2022-02-23 writing/appendix_a.Rmd
Removed sections duplicated between paper, supplementay info, and appendix

---
title: "Supplementary Information"
short_title: COVID-19 Prescribing
bibliography:
- r-references.bib
- ../references/library.bib
output:
  pdf_document: default

papersize: a4
header-includes:
  \usepackage{float}
  \let\origfigure\figure
  \let\endorigfigure\endfigure
  \renewenvironment{figure}[1][2] {
      \expandafter\origfigure\expandafter[H]
  } {
      \endorigfigure
  }

---

```{r setup, include = FALSE}
## Packages
# devtools::install_github("w-hardy/serCymruTools")
library(plyr); library(dplyr)
library(rms)
library(nlme)
library(tidyverse)
library(lubridate)
library(tsibble)
library(fable)
library(feasts)
library(furrr)
library(rebus)
library(serCymruTools)
library(knitr)
library(kableExtra)
library(papaja)
library(fuzzyjoin)
library(ggbiplot)

r_refs("r-references.bib")

## Data
gp <- 
  read_rds("../data/gp_agesex_qof_combined.rds") %>% 
  mutate(npat = N,
         N = NULL,
         practice_id = as.character(practiceid),
         hboard = case_when(hboard == "W11000023" ~ "BC",
                            hboard == "W11000024" ~ "P",
                            hboard == "W11000025" ~ "HD",
                            hboard == "W11000028" ~ "AB",
                            hboard == "W11000029" ~ "CaV",
                            hboard == "W11000030" ~ "CTM",
                            hboard == "W11000031" ~ "SB")) %>% 
  ungroup() %>% 
  janitor::clean_names() %>% 
  mutate(dispensing = case_when(tolower(dispensing) == "yes" ~ TRUE, 
                                tolower(dispensing) == "no" ~ FALSE),
         hboard = as.factor(hboard))

# GP extract data
data_files <-
  paste0("../data/prescribing_data/", list.files("../data/prescribing_data")) 
names(data_files) <- 
  str_remove_all(list.files("../data/prescribing_data"), ".rds")
pres_data <- 
  map(data_files, read_rds)

# CV model accuracies
cv_model_files <- 
  paste0("../data/gp_cv_models/", 
         list.files("../data/gp_cv_models/"))
names(cv_model_files) <- 
  cv_model_files %>% 
  str_remove_all("../data/gp_cv_models/") %>% 
  str_remove_all("_gp_cv_models.rds")

# Forecast - observed
differences_files <- 
  list.files("../data/differences/")
names(differences_files) <- 
  str_remove_all(differences_files, "_discreps.rds")

plan(cluster)
differences <- 
  furrr::future_map_dfr(.x = differences_files,
                        .f = ~read_rds(paste0("../data/differences/", .x)),
                        .id = "drug") %>% 
  mutate(drug = relevel(as_factor(drug), ref = "quant_all_drugs"),
         drug = str_remove_all(drug, "ddd_"),
         drug = str_remove_all(drug, "quant_"),
         drug = str_remove_all(drug, "_ddd")) %>% 
  filter(regional_unit != "W95636") #https://abuhb.nhs.wales/news/news/important-information-regarding-lansbury-surgery-caerphilly/

plan(sequential)

drugs <- unique(differences$drug)

# Seed for random number generation
set.seed(32)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE, 
                      echo = FALSE, fig.pos = 'h!')
kable_opts <- options(knitr.kable.NA = "-")
```

\renewcommand{\appendixname}{Supplementary Material}
\renewcommand{\thefigure}{SI\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{SI\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{SI\arabic{table}} \setcounter{equation}{0}

## GLS Results

### Analysis

```{r, gls-functions}

#### GLS Assumption Checking ####
glsAssump <- function(model, data){
  data$resid <- r <- resid(model); data$fitted <- fitted(model)
  y1 <- ylab("Residuals")
  p1 <- ggplot(data, aes(x = fitted, y = resid)) + geom_point(alpha = .1) + y1
  # p2 <- ggplot(data, aes(x = feb_diff, y = resid)) + geom_point(alpha = .1) + y1
  p3 <- ggplot(data, aes(x = Month, y = resid)) + y1 + 
    stat_summary(fun.data = "mean_sdl", geom = "smooth")
  p4 <- ggplot(data, aes(sample = resid)) + stat_qq(alpha = .1) + 
    geom_abline(intercept = mean(r), slope = sd(r)) + y1

  gridExtra::grid.arrange(p1, #p2, 
                          p3, p4, ncol = 2)
}

#### GLS Estimated Effects ####

glsEffects <- function(model, name, data){

  Name <- case_when(name == "all_drugs" ~ "All",
                    name == "ace" ~ "ACEi",
                    name == "adreno" ~ "BAA",
                    name == "cortico" ~ "Inhaled \ncortico",
                    name == "hcq" ~ "HCQ",
                    name == "nsaid" ~ "NSAIDs",
                    name == "oral_contra" ~ "Oral\ncontra",
                    name == "paracet" ~ "Paracet",
                    name == "ssri" ~ "SSRIs",
                    name == "vitd" ~ "Vit D",
                    name == "warfarin" ~ "Warfarin")

  cat("\n\n")

  #### Plot interactions #### not including month*ngp as none significant
  p1 <- Predict(model, Month, dispensing, hboard, conf.int = .95) %>% 
    predictPlot(., groups = hboard, "Health board")
  gridExtra::grid.arrange(p1,nrow = 1, newpage = TRUE) # Effects 1
  cat("\n\n")

  p2 <- Predict(model, Month,  dispensing, hboard, conf.int = .95) %>% 
    predictPlot(., groups = dispensing, "Dispensing")
  gridExtra::grid.arrange(p2,nrow = 1, newpage = TRUE) # Effects 2
  cat("\n\n")

  p1 <- Predict(model, Month, dispensing, hboard,
                wimd2019 = percentiles(data, wimd2019),
                conf.int = .95) %>% 
        group_by(Month, wimd2019) %>% 
    summarise(yhat = mean(yhat)) %>% 
    predictPlot(., groups = wimd2019, "WIMD")
    gridExtra::grid.arrange(p1, nrow = 1, newpage = TRUE) # Effects 3
  cat("\n\n")

  p1 <- Predict(model, Month, dispensing, hboard,
                ngp = percentiles(data, ngp),
                conf.int = .95) %>% 
    predictPlot(., groups = ngp, "Number of GPs")
  gridExtra::grid.arrange(p1, ncol = 1, newpage = TRUE) # Effects 5
  cat("\n\n")

  p2 <- Predict(model, Month, dispensing, hboard,
                gp_load = percentiles(data, gp_load),
                conf.int = .95) %>% 
    predictPlot(., groups = gp_load, "GP load")
  gridExtra::grid.arrange(p2, nrow = 1, newpage = TRUE) # Effects 4
  cat("\n\n")

  # p1 <- Predict(model, Month, dispensing, hboard,
  #               PC1 = percentiles(data, PC1),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC1, "PC1")
  # p2 <- Predict(model, Month, dispensing, hboard,
  #               PC2 = percentiles(data, PC2),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC2, "PC2")
  # p3 <- Predict(model, Month, dispensing, hboard,
  #               PC3 = percentiles(data, PC3),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC3, "PC3")
  # p4 <- Predict(model, Month, dispensing, hboard,
  #               PC4 = percentiles(data, PC4),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC4, "PC4")
  # gridExtra::grid.arrange(p1,p2,p3,p4, ncol = 2, newpage = TRUE) # Effects 6
  # cat("\n\n")

  # #### ANOVA plot ####
  # p <- 
  #   capture.output(anova(model))[4:41] %>%
  #   read_fwf(fwf_empty(., col_names = c("Factor", "Chi-Square", "d.f.", "P"))) %>% 
  #   mutate(Factor = 
  #            str_remove_all(Factor, 
  #                           c(OPEN_PAREN %R% "Factor" %R% PLUS %R% 
  #                               "Higher Order Factors" %R% CLOSE_PAREN))) %>% 
  #   filter(!str_detect(Factor, "Nonlinear") &
  #            !str_detect(Factor, "TOTAL") &
  #            Factor != "All Interactions",
  #          !str_detect(Factor, c("f" %R% OPEN_PAREN %R% "A,B"))) %>% 
  #   ggplot(aes(x = `Chi-Square`-`d.f.`, 
  #              y = reorder(Factor, X = `Chi-Square`-`d.f.`))) +
  #   geom_point(aes(col = 
  #                    as.factor(if_else(as.numeric(str_remove_all(P, "<"))<.05, 
  #                                      TRUE,FALSE)))) +
  #   scale_color_discrete(name = "P < .05") + 
  #   xlab(expression(chi^2-"d.f.")) +
  #   ylab(NULL) +
  #   ggthemes::theme_few() 
  # 
  # gridExtra::grid.arrange(p, newpage = TRUE)
  # cat("\n\n")

  p <- anova(model, rm.other = list("Month"), trans = sqrt,
             tol = 1e-10, dec.chisq = 2, dec.p = 2) %>% 
    plot()

  p

  #### ANOVA Table ####
  # model %>% 
  #   anova() %>% 
  #   kable(booktabs = TRUE, linesep = "", digits = 2,
  #         caption = paste(Name, "Wald z-test statistics")) %>% 
  #   # add_indent(positions = c(3,5,6,8,10,11,13,15,17,19,21,23,25,27,28,30:34)) %>% 
  #   kable_styling(latex_options = c("striped", "HOLD_position", "repeat_header")) %>% 
  #   print()

  cat("\n\n")
}

```

```{r, gls-analysis-read}
#### gls-analysis-read ####
best_models <- read_rds("../data/gls/best_models.rds")
best_bs_models <- read_rds("../data/gls/best_bs_models.rds")
```

```{r, gls-effects, results="asis", fig.cap=gls_analysis_fig_cap, dependson="gls-functions"}
#### gls-effects ####

dd <- NULL
for(i in 1:length(best_bs_models)){
  cat("####", (drugs[i]), "\n\n")
  cat("#####", "Assumptions\n\n")
  gls_data <- glsData(drugs[i])
  dd <- datadist(gls_data)
  options(datadist = "dd")
  glsAssump(best_bs_models[[i]], gls_data)
  cat("\n\n")
  cat("#####", "Effects\n\n")
  glsEffects(best_bs_models[[i]], drugs[i], gls_data)
  rm(dd)
  cat("\n\n")
  options(datadist = NULL)
}

```

### Summary of effects

```{r, gls-anova-summary-tab}
#### gls-anova-summary-tab ####
best_bs_models %>% 
  map(~anova(., tol = 1e-10)) %>% 
  map(~as_tibble(x = ., rownames = "Coef")) %>% 
  map(as.matrix) %>% 
  map(as_tibble) %>% 
  bind_rows(.id= "drug") %>% 
  mutate(across(`Chi-Square`:P, as.numeric)) %>% 
  select(-c(`Chi-Square`, d.f.)) %>% 
  filter(!str_detect(Coef, "Nonlinear") &
           !str_detect(Coef, "TOTAL") &
           !str_detect(Coef,"All Interactions"),
         !str_detect(Coef, c("f" %R% OPEN_PAREN %R% "A,B"))) %>%
  pivot_wider(names_from = drug, values_from = P) %>% 
  mutate(across(ace:warfarin, ~printnum(., zero = FALSE)),
         Coef = str_remove_all(Coef, OPEN_PAREN %R% "Factor" %R% PLUS %R% 
                                 "Higher Order Factors" %R% CLOSE_PAREN)) %>% 
  select(Coef, all_drugs, everything()) %>% 
  kable(booktabs = TRUE, linesep = "",
        caption = paste("Wald test p-values")) %>% 
  kable_styling(latex_options = c("striped", "HOLD_position", "repeat_header")) %>% 
  landscape()
```

### Full tables

```{r, gls-full-model-table, results="asis"}
options(prType="latex")

best_bs_models %>% 
  map(~print(., latex=TRUE))

options(prType="plain")
```

\newpage

## SM References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage